### 2014-04-25 sequential/latex/DAC2013/theory.tex

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{acm_proc_article-sp}

%packages
\usepackage[linesnumbered, ruled]{algorithm2e}
\RequirePackage{amssymb, mathptm}
\usepackage{amsbsy}
\usepackage{graphicx}
\usepackage{helvet}
\usepackage{enumerate}
\usepackage{amsmath}
\usepackage{cases}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfig}
\usepackage{comment}

%new format
\newtheorem{Algorithm}{Algorithm}
\newdef{Definition}{Definition}
\newtheorem{Theorem}{Theorem}
\newtheorem{Lemma}{Lemma}
\newdef{Example}{Example}

\begin{document}

\title{Verification of Sequential Galois Field Circuits using Algebraic Geometry
\titlenote{Version 0.3, refined theory part with normal basis, proposed details for designing 2 examples.}}
%\subtitle{[Extended Abstract]
%\titlenote{A full version of this paper is available as
%\textit{Author's Guide to Preparing ACM SIG Proceedings Using
%\LaTeX$2_\epsilon$\ and BibTeX} at
%\texttt{www.acm.org/eaddress.htm}}}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{2} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Xiaojun Sun\\
       \affaddr{University of Utah}\\
       \affaddr{Department of Electrical \& Computer Engineering}\\
       \affaddr{Salt Lake City, USA}\\
       \email{xiaojun.sun@utah.edu}
% 2nd. author
\alignauthor
Priyank Kalla\\
       \affaddr{University of Utah}\\
       \affaddr{Department of Electrical \& Computer Engineering}\\
       \affaddr{Salt Lake City, USA}\\
       \email{kalla@ece.utah.edu}
% 3rd. author
%\alignauthor Lars Th{\o}rv{\"a}ld\titlenote{This author is the
%one who did all the really hard work.}\\
%       \affaddr{The Th{\o}rv{\"a}ld Group}\\
%       \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
%       \affaddr{Hekla, Iceland}\\
%       \email{larst@affiliation.org}
%\and  % use '\and' if you need 'another row' of author names
% 4th. author
%\alignauthor Lawrence P. Leipuner\\
%       \affaddr{Brookhaven Laboratories}\\
%       \affaddr{Brookhaven National Lab}\\
%       \affaddr{P.O. Box 5000}\\
%       \email{lleipuner@researchlabs.org}
% 5th. author
%\alignauthor Sean Fogarty\\
%       \affaddr{NASA Ames Research Center}\\
%       \affaddr{Moffett Field}\\
%       \affaddr{California 94035}\\
%       \email{fogartys@amesres.org}
% 6th. author
%\alignauthor Charles Palmer\\
%       \affaddr{Palmer Research Laboratories}\\
%      \affaddr{8600 Datapoint Drive}\\
%%      \affaddr{San Antonio, Texas 78229}\\
%       \email{cpalmer@prl.com}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{21 Jan 2014}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle
\begin{abstract}
Circuits working in Galois fields are increasingly employed in designs like arithmetic component for Elliptic Curve Cryptography (ECC). 
This work proposes a new method to effectively verify sequential circuits in Galois fields. Algebraic geometry
is introduced to describe circuits behavior and expand the definition of implicit state space traversal. Moreover, Gr\"obner basis
representation is adopted for word-level abstraction on circuit variables to address state explosion problem
with BDDs. Experiments are run with fast Gr\"obner basis computation engine to get competitive results.
\end{abstract}

% A category with the (minimum) three required fields
\category{EDA5.1}{Verification}{Functional, transaction-level, RTL, and gate-level modeling and verification of hardware design}
%A category including the fourth, optional field follows...
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

\terms{Verification, Algorithms}

\keywords{Verification, Sequential, Galois Fields, Algebraic Geometry} % NOT required for Proceedings

\section{Essential Theory}
\subsection{Polynomial Representation}
A typical sequential circuit is composed as figure \ref{fig:sequential}. The combinational logic has primary input $x$, state
inputs $\{s_0, s_1, \dots, s_{k-1}\}$, primary output $z$ and state outputs $t_0, t_1, \dots, t_{k-1}$. State outputs
are latched into registers, which will be fed back to state inputs in next clock cycle.

\begin{figure}
\centering
\psfig{file=./sequential_fig.ps, height=2in, width=3.5in,}
\caption{A typical sequential circuit block view}
\label{fig:sequential}
\end{figure}

A transition function $\Delta$ is a Boolean function about state outputs and all inputs, i.e.
\begin{displaymath}
t_i = \Delta_i(x, s_0, s_1, \dots, s_{k-1})
\end{displaymath}
This constrain could also be written in the form of equation $polynomial = 0$:
\begin{displaymath}
t_i - \Delta_i(x, s_0, s_1, \dots, s_{k-1}) = 0
\end{displaymath}
Solution to this equation is limited within $\{0, 1\}$, which is set of all elements in Galois field $\mathbb{F}_2$.
Furthermore, there exists a one-to-one corresponding relation from Boolean space to Galois field on higher dimension: 
$\mathbb{B}^k \to \mathbb{F}_{2^k}$, which means it is possible to find a set of elements in $\mathbb{F}_{2^k}$ representing
all Boolean values every bit of a bit vector can take. In Galois field, a as \ref{table:booltogalois} shows.
\begin{table}[h]
\centering
\caption{Bit-vector, Exponential and Polynomial representation of
elements in  ${\mathbb{F}}_{2^4} = {\mathbb{F}}_2[x]
\pmod{x^4+x^3+1}$}\label{table:booltogalois}  
\begin{tabular}{|c|c||c|c|} 
\hline
$a_3a_2a_1a_0$ & Polynomial     &$a_3a_2a_1a_0$ & Polynomial  \\
\hline
$0000$        & $0$           & $1000$  &$\alpha^3$\\
\hline
$0001$        & $1$           & $1001$  & $\alpha^3 + 1$\\
\hline
$0010$        &  $\alpha$       & $1010$ & $\alpha^3 + \alpha$  \\
\hline
$0011$        &  $\alpha + 1$   & $1011$ &  $\alpha^3+\alpha+1$\\
\hline
$0100$        &  $\alpha^2$     &  $1100$ &  $\alpha^3 + \alpha^2$\\
\hline
$0101$        & $\alpha^2 + 1$ & $1101$  & $\alpha^3+\alpha^2+1$\\
\hline
$0110$        &  $\alpha^2 + \alpha$ & $1110$ &  $\alpha^3+\alpha^2+\alpha$\\
\hline
$0111$        & $\alpha^2+\alpha+1$ & $1111$ & $\alpha^3+\alpha^2+\alpha+1$\\
\hline
\end{tabular}
\end{table}
A word-level representation on state inputs/outputs can be attained through above correspondence. If the set of 
state inputs $\{s_0, s_1, \dots, s_{k-1}\}$ is denoted by $S$ taking values from $\mathbb{F}_{2^k}$, and
set of outputs $\{t_0, t_1, \dots, t_{k-1}\}$ is denoted by $T$, a word-level transition function can be written as
\begin{displaymath}
T - \Delta(x, S) = 0
\end{displaymath}
If $T - \Delta(x, S)$ is taken as a polynomial $f$, then $f \in \mathbb{F}_{2^k}[x]\ mod\ P(\alpha)$.
\subsection{Gr\"obner Basis Theory}
Gr\"obner basis is used to calculate desired results out of a set of polynomials.
If there exists multiple polynomials $\{f_1, f_2, \dots, f_r\}$ representing constrains on $T, x$ and $S$,
then all polynomials should simultaneously equal to $0$.
\begin{displaymath}
\left\{
  \begin{array}{lc}
  f_1 & = 0\\
  f_2 & = 0\\
  \dots & \ \\
  f_r & = 0
  \end{array} \right.
\end{displaymath}
Solution to these equations is set of values of $(T, x, S)$, which is called \textbf{Variety} of \textbf{ideal}
$I = \langle f_1, f_2, \dots, f_r\rangle $.

Calculate Reduced Gr\"obner Basis (RGB) from ideal $I$ with \textbf{elimination term order}, there exists a polynomial contains
only $T$, if the values of states input $S$ and primary input $x$ are given in the form of polynomials and included by $I$,
the values of $T$ can be computed.

Otherwise, if RGB calculation is manipulated under \textbf{abstraction term order}, there exists a polynomial in
the form of $T - \mathcal{F}(x, S)$ representing transition function. In next clock cycle, assign new state input $S'$ with $T$, the next
state can be computed again, result is polynomial $T' - \mathcal{F}(x', S')$.

\subsection{Normal Basis Representation}

Let $\beta$ be an element in the Galois field $F_{2^n}$ constructed by primitive element $\alpha$ and minimal polynomial
$P(\alpha)$. Then a basis in the form $\{\beta, \beta^2, \beta^4, \beta^8, ... ,\beta^{2^{n-1}}\}$ is a
\emph{Normal Basis}; here $\beta$ is called \emph{Normal Element}.

For arithmetic operations in Galois fields, squaring and multiplication (with modulus) can be greatly simplified if 
normal basis representation is adopted to represent operands.
\begin{Example}
\label{ex:nb_sq}
Element squaring: In $F_{2^n}$, all coefficients which can be divided by 2 are reduced, so 
following equality holds for all field elements $a$ and $b$:
$$(a+b)^2 = a^2 + b^2$$ 
Apply this rule on element squaring of standard/polynomial basis:
\begin{align}
& (b_0\beta + b_1\beta^2 + b_2\beta^4 + \dots + b_{n-1}\beta^{2^{n-1}})^2 \nonumber\\
&= b_0^2\beta^2 + b_1^2\beta^4 + b_2^2\beta^8 + \dots + b_{n-1}^2\beta^{2^n} \nonumber\\
&= b_{n-1}^2\beta + b_0\beta^2 + b_1\beta^4 + \dots + b_{n-2}\beta^{2^{n-1}} \nonumber
\end{align}
using Fermat's little theorem $\beta^{2^n} = \beta$. This shows that squaring of field elements
represented by normal bases can be easily completed by operating a simple right-cyclic rotation.
\end{Example}

\begin{Example}
\label{ex:nb_multi}
Normal basis multiplication: There are 2 binary vectors representing operands of multiplication in normal
basis: $$A = (a_0, a_1, \dots, a_{n-1}),\ B = (b_0, b_1, \dots, b_{n-1})$$ 
the product is also written in normal basis representation: $$C = A*B = (c_0, c_1, \dots, c_{n-1})$$
Describe calculation procedure for the most significant digit of product with a function: 
$$c_{n-1} = f(a_0, a_1, \dots, a_{n-1}; b_0, b_1, \dots, b_{n-1})$$
Square both side: $C^2 = A^2*B^2$, i.e. the second significant digit 
$$c_{n-2} = f(a_{n-1}, a_0, a_1, \dots, a_{n-2}; b_{n-1}, b_0, b_1, \dots, b_{n-2})$$ 
By this method it is easy to get all digits of product $C$.
\end{Example}

\section{Algebraic Geometry Approach}
\subsection{State Space Traversal}

\begin{figure}
\centering
\psfig{file=./fsm_fig.ps, height=2in, width=3.5in,}
\caption{Gate-level circuit of combinational logic block of sample FSM}
\label{fig:fsm}
\end{figure}

\begin{figure}
\centering
\psfig{file=./stg_fig.ps, height=2in, width=3.5in,}
\caption{State transition graph of sample FSM}
\label{fig:stg}
\end{figure}

The first example is a mealy finite state machine (FSM). Gate-level circuits of combinational logic
is figure \ref{fig:fsm}, $x$ is primary input,
$\{s_0, s_1\}$ are state inputs, $Z$ is primary output and $\{t_0, t_1\}$ are state outputs.
Word-level state input is $S = s_0 + s_1\cdot\alpha$, output is $T = t_0 + t_1\cdot\alpha$. In Galois field
$\mathbb{F}_{2^2}$, bit-level variables $x, s_0, s_1, Z, t_0, t_1$ can take values from $\{0, 1\}$, while
word-level variables $S, T$ may take values from $\{0, 1, \alpha, 1 + \alpha\}$. State transition graph (STG)
showed in figure \ref{fig:stg} uses 2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2, S_3\}$, which
could be converted to elements in $\mathbb{F}_{2^2}$ similarly as table \ref{table:booltogalois} shows.

One important technique to check sequential equivalence is \textbf{state space traversal}. In this example,
Breath-First-Search (BFS) is employed for traversal, which can be described with following algorithm:

\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}

  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}\label{alg:BFS}
\end{algorithm}

Our approach is implementing state space traversal by refined BFS algorithm combined with polynomial representation
and Gr\"obner basis theory. In practice, this example will show how to apply concepts and techniques from algebraic
geometry on each line of above BFS algorithm.

\subsection{States and Varieties of Ideal}

\begin{Theorem}
State variables $S, T$ and sets of states such as $from^i, to^i$ can always be represented as varieties of ideals.
\end{Theorem}
For example in Line 1 of BFS algorithm, assume initial state is $S_3$ in figure \ref{fig:stg}, then corresponding
polynomial $f = \mathcal{F}(S^0) = S^0 - 1 - \alpha$. Consider an ideal $I$ with only one generator $f$, its variety
$V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma - 1 - \alpha = 0\}$, which equals to $\{1 + \alpha\}$, the only
valid value $S^0$ can take.

If an ideal contains multiple polynomial specifications, it is necessary to compute Gr\"obner basis with elimination term
order to get one polynomial only on desired variable. In first iteration of Line 4, $to^i$ has multiple specifications,
some of them are transition functions on bit-level:
\begin{displaymath}
\begin{array}{ll}
f_1:& t_0 - (\overline{x}\ and\ \overline{s_0}\ and\ \overline{s_1})\ or\ (s_0\ and\ s_1) \\
f_2:& t_1 - (\overline{s_0}\ and\ x)\ or\ (p\ and\ \overline{s_1})\
\end{array}
\end{displaymath}
And some are bits-to-word definitions fitting polynomial representation of elements in $\mathbb{F}_{2^2}$:
\begin{displaymath}
\begin{array}{ll}
f_3:& S - s_0 - s_1\alpha \\
f_4:& T - t_0 - t_1\alpha
\end{array}
\end{displaymath}
And an polynomial about initial state as mentioned above:
\begin{displaymath}
f_5:\  S - 1 - \alpha
\end{displaymath}
And the rests are vanishing polynomials for every variable, bit-level and word level:
$f_6: x^2 - x; f_7: t_0^2 - t_0; f_8: t_1^2 - t_1; f_9: S^4 - S; f_{10}: s_0^2 - s_0;
f_{11}: s_1^2 - s_1; f_{12}: T^4 - T$

Transition equations here contains some Boolean operators, they can be re-written in terms of operations in Galois fields. 
In $\mathbb{F}_{2}$, let $TRUE = 1, FALSE = 0$, for either element $a$ in this field, considering $0 + 0 = 1 + 1 \equiv 0\ (mod\ 2)$
and $0 + 1 = 1 + 0 \equiv 1\ (mod\ 2)$, the inverse of $a$ is: $\overline{a} = 1 + a$. Similarly all Boolean operators
can be converted within $\mathbb{F}_{2}$, table \ref{table:booltogalois_op} gives part of them and their corresponding
operations in $\mathbb{F}_{2}$:
Also note that there is no specification on initial primary input $x$, in Gr\"obner basis approach this means $x$ is smoothed by
reversely using Shannon's expansion.
\begin{table}
\centering
\caption{Some Boolean operators and corresponding operations in $\mathbb{F}_{2}$}
\label{table:booltogalois_op}
\begin{tabular}{|c|c|} \hline
Boolean operator & operation in $\mathbb{F}_{2}$\\ \hline
$\overline{a}$ & $1 + a$\\ \hline
$a\ and\ b$ & $ab$\\ \hline
$a\ or\ b$ & $a + b + ab$\\ \hline
$a \oplus b$ & $a + b$\\
\hline\end{tabular}
\end{table}
Using elimination term order: $others\ >\ bit$-$level\ inputs/outputs\ >\ S\ >\ T$, compute Gr\"obner basis for ideal
$J = \langle f_1, f_2, \dots, f_{12}\rangle $, the result will include one polynomial $f_T$ contains only variable $T$. In this example,
$f_T = T + 1$. This polynomial equals to $T - 1$ since coefficients of polynomial representation in $\mathbb{F}_{2^2}$ are
limited within $\mathbb{F}_{2}$, where $1 \equiv -1\ (mod\ 2)$. Solution to $f_T = 0$ is $T = 1$, which shows that next
state the machine just reached is $\{S_1(01)\}$.

\subsection{Application of Algebraic Geometry}

There are some difficulties with polynomial representation when executing Line 5 and Line 6, it is necessary to explore
how \textbf{Union}, \textbf{Intersection} and \textbf{Complement Set} works in Galois field $\mathbb{F}_{2^k}$. Since
state set variables $from^i, to^i, reached$ all have single polynomial, consider an ideal $I$ with single generator $I = \langle f\rangle $.
Example: assume $I = \langle f\rangle  = \langle T^2 + (1+\alpha)\cdot T+\alpha\rangle $, its variety $V(I) = \{a\ |\ a \in \mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$.
So it is reasonable to specify: union, intersection and complement set mentioned in this paper are all functions on \textbf{varieties}.
To better discuss this problem, introduce following concepts from algebraic geometry:
\begin{Definition}
\label{def:sum}
({\bf Sum of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the 
{\bf sum} of $I$ and $J$, denoted by $I + J$, is the set
  \begin{equation}
  I + J = \{f + g\ |\ f \in I \ and\  g \in J\}.
  \end{equation}
Furthermore, if $I = \langle f_1, \dots, f_r\rangle$ and 
$J = \langle g_1, \dots, g_s\rangle$, then 
$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$.
\end{Definition}
\begin{Definition}
\label{def:prod}
({\bf Product of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the
{\bf product} of $I$ and $J$, denoted by $I \cdot J$, is defined to be the ideal generated 
by all polynomials $f \cdot g$ where $f \in I$ and $g \in J$. Furthermore, let
$I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$, then
  \begin{equation}
  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle .
  \end{equation}
\end{Definition}
\begin{Definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then $I:J$
is the set
  \begin{equation}
  \{f \in k[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{Definition}
These concepts can lead the way to solution by adopting following theorems:
\begin{Theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then ${\bf V}(I + J) = {\bf V}(I)
\bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) = {\bf V}(I) \bigcup {\bf V}(J)$.
\end{Theorem}
\begin{Theorem}
\label{thm:quotient}
If $I, J$ are ideals with only one generator, then ${\bf V}(I:J) ={\bf V}(I) - {\bf V}(J)$.
\end{Theorem}
Proof to these theorems are referred to (that yellow book?) and (my write-up?).

For varieties intersection $\{1\}\bigcap\{1, \alpha\}$, calculate ideal sum $\langle T+1, T^2 + (1+\alpha)\cdot T+\alpha\rangle  = \langle T+1\rangle $,
its variety is $\{1\}$; for varieties union $\{1,1+\alpha\}\bigcup\{1+\alpha\}$, calculate
ideal product $\langle (T+1+\alpha)\cdot(T^2 + (1+\alpha)\cdot T+\alpha)\rangle  = \langle T^3 + 1\rangle $, its variety
is $\{1, \alpha, 1+\alpha\}$; for complement set of variety $\{1, \alpha\}$, the universal set is
the variety of ideal of vanishing polynomial $V(\langle T^4-T\rangle ) = \{0,1,\alpha,1+\alpha\}$,
so $\overline{V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )} = V(\langle T^4-T\rangle ) - V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )$,
which equals to $V(\langle T^4-T\rangle :\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle ) = V(\langle T^2+(1+\alpha)\cdot T\rangle )$,
the result is $\{0,1+\alpha\}$.

From discussion above, the BFS traversal is completely implemented in Galois field.
If the initial input is $S_3$: $S + 1 + \alpha$, the final return value will be set of
reachable states: $T^4 + T$, i.e. the universal set, $\{S_0, S_1, S_2, S_3\}$.

\subsection{Discussion on Ideal quotient and its generator}
In nature, we are looking varieties as the solutions to specific polynomial $f=0$ WITHIN $\mathbb{F}_{2^n}$, because (affine) varieties are defined
within algebraic closure $k^n$ but we only care about those fall in $\mathbb{F}_{2^n}$.

Theorem \ref{thm:unionintersect} related to DEF. \ref{def:sum} and \ref{def:prod} counts on ideal generators, so there will be no ambiguity on 
new ideals we get from this theorem. It also appeared in Avrunin's CAV paper. The only problem is what ideal quotient operation will give us.
This should be further discussed because the completeness of our algebraic geometry approach will benefit from it. \emph{Singular} has the function of
ideal quotient, so ask from the author maybe another choice.

If we put this conception aside, it is also feasible to define complement set from "solution to 
equation" sense. Assume $f(x) = x + 1 + \alpha = 0$, its solution within $\mathbb{F}_{2^2}$ 
is $\{1+\alpha\}$; meanwhile solutions to vanishing polynomial $v(x) = x^2 + x = 0$ are
$\{0,1,\alpha,1+\alpha\}$. Now compute solutions to $$g(x) = \frac{v(x)}{f(x)} = 0$$,
this requires $v(x) = 0$ and $f(x) \neq 0$, which means the complement set of solution 
to $f(x) = 0$.

Issues to this interpretation: (a) Is $v(x)$ always divisible by $f(x)$? (b) Is $v(x)=0$ and
$f(x) \neq 0$ equivalent with $g(x)=0$ on their solutions? Could $f(x)$ equal to $0$?

\section{Functional Verification}
\subsection{Sequential Galois Arithmetic Multiplier}
Our experiment is to verify the function of a Galois field multiplier after certain clock cycles.
The following design is Sequential Multiplier with Parallel Output (SMPO),
a Normal Basis multiplier based on Massey-Omura algorithm. Inputs and outputs
are all 5-bit word taking value from ${\mathbb F}_{2^5}$. After loading operands
 A and B, setting all output latches to 0 and running for 5 iterations, the 
output should be $R = A\cdot B$ (mod $x^5 + x^2 + 1$).
\begin{figure}
\centering
\epsfig{file=./mySMPO.eps, height=2in, width=3.5in}
\caption{5-bit SMPO}
\label{fig:SMPO}
\end{figure}
Similarly, build elimination ideal for all gates/operations and induce
initial states of latches. However, instead of eliminating all variables
to one, this example adopts abstraction term order
and keeps the polynomial contains function between output and inputs.
Here the lex ordering is $\{others\} > R > \{A, B\}$, and objective
polynomial is $R + \mathcal{F}(A, B)$.

Apply this approach to calculate image in BFS traversal, but modify algorithm
\ref{alg:BFS} to make it adapt 5 steps reachable states enumeration.
The result is $R + AB$, which validates the function of this circuit.

\subsection{Gr\"obner basis based Approach}

For 5-bit normal basis multiplication, the $i$-th digit of product can be written as
\begin{align}
R_i &= b_ia_{i+1} + b_{i+1}(a_i + a_{i+3}) + b_{i+2}(a_{i+3} + a_{i+4}) \nonumber\\
&+ b_{i+3}(a_{i+1} + a_{i+2}) + b_{i+4}
(a_{i+2} + a_{i+4}),\ 0\leq i\leq 4 \nonumber
\end{align}
It is possible to calculate every conjunctive term in the product simultaneously within one clock cycle,
on distinct bits. 
Use cyclic similarity of above function, following executions are operated: 
\begin{Example}
Sequential Multiplier Protocol:
\begin{itemize}
\item \textbf{Initial}\ \ $R_0 = R_1 = R_2 = R_3 = R_4 = 0$
\item \textbf{Clock 1}\ \ $R_0 = a_1b_0, R_1 = b_2(a_1 + a_4), R_2 = b_4(a_0 + a_1), R_3 = b_1(a_4 + a_0), 
			R_4 = b_3(a_1 + a_3)$
\item \textbf{Clock 2}\ \ $R_0 = b_3(a_1 + a_3) + a_0b_4, R_1 = a_1b_0 + b_1(a_0 + a_3), R_2 = b_2(a_1 + a_4)
			+ b_3(a_4 + a_0), R_3 = b_4(a_0 + a_1) + b_0(a_3 + a_4), R_4 = b_1(a_4 + a_0) + b_2(a_0 + a_2)$
\item \textbf{$\dots$}
\item \textbf{Clock 5}\ \ $R_0 = c_0, R_1 = c_1, R_2 = c_2, R_3 = c_3, R_4 = c_4$, i.e. $R = A\cdot B$.
\end{itemize}
\end{Example}

In BFS algorithm, the \textbf{Image} function is implemented by constructing an elimination ideal then 
compute its Gr\"obner basis. 
\begin{Example}
\label{ex:SMPO}
For 5-bit SMPO, the ideal consists of (for the first clock cycle):

(a) {\bf Gate descriptions:}
$a_1+a_4+c_1, a_1+a_0+c_2, a_0+a_4+c_3, a_1+a_3+c_4,
		  a_1b_0+r_4+R_0, c_1b_2+r_0+R_1, c_2b_4+r_1+R_2, c_3b_1+r_2+R_3, c_4b_3+r_3+R_4;$

(b) {\bf Word-level variables:}
$A+a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18},
		  B+b_0\alpha^5+b_1\alpha^{10}+b_2\alpha^{20}+b_3\alpha^9+b_4\alpha^{18},
		  r+r_0\alpha^5+r_1\alpha^{10}+r_2\alpha^{20}+r_3\alpha^9+r_4\alpha^{18},
		  R+R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18};$

(c) {\bf Vanishing polynomials:}
		 $ a_0^2+a_0, a_1^2+a_1, a_2^2+a_2, a_3^2+a_3, a_4^2+a_4,
		  b_0^2+b_0, b_1^2+b_1, b_2^2+b_2, b_3^2+b_3, b_4^2+b_4,
		  r_0^2+r_0, r_1^2+r_1, r_2^2+r_2, r_3^2+r_3, r_4^2+r_4,
		  R_0^2+R_0, R_1^2+R_1, R_2^2+R_2, R_3^2+R_3, R_4^2+R_4,
		  c_1^2+c_1, c_2^2+c_2, c_3^2+c_3, c_4^2+c_4,
		  A^{32}+A, B^{32}+B, r^{32}+r, R^{32}+R;$

(d)	{\bf Feedback input:}	  $r_{in}$.

Polynomial $r_{in}$ equals to $from^{i-1}$ in Line 4, BFS algorithm. Under abstraction term ordering,
polynomial $to^i$ is assigned with a polynomial in result Gr\"obner basis which has the form $R + \mathcal{F}(A,B)$. 
Since all outputs are connected to feedback inputs
in SMPO, Line 5 and 6 will be neglected. Line 7 is finished by replace current output $R$ with previous 
output $r$. Initially $r_{in} = 0$.

In next clock cycle, $r_{in} = r + \mathcal{F}(A,B)$ updated from latest result; simultaneously operands
$A$ and $B$ have been cyclically shifted, so gate descriptions in (a) should be modified accordingly. The 
loop is operated for 5 times, result from the last step's Gr\"obner basis should be polynomial $R + AB$.
\end{Example}

This experiment can be repeated on $n$-bits SMPO after running for $n$ clock cycles.

\section{Fast Gr\"obner basis computation}
\subsection{Refined abstraction term order (RATO)}
A lexicographic order constrained by following relation $>_{r}$: "circuit variables ordered reverse topologically" $>$ 
"designated word-level output" $>$ "word-level inputs" is called the \emph{Refined Abstraction Term Order (RATO)}.

In Buchberger's algorithm, the specification polynomial (Spoly) for each pair is calculated. In RATO, most polynomials
have relatively prime leading terms/monomials (which means $Spoly \xrightarrow{J+J_0}_{+} 0$) except one pair:
word-level polynomial corresponding to outputs and its leading bit-level variable's gate description polynomial.
Its remainder $r$ lets following lemma hold:

\begin{Lemma}
\label{lem:1}
$r$ will only contain primary inputs (bit-level and word-level) and word-level output; furthermore, there will be one and
only one term containing word-level output whose monomial is word-level output itself rather than higher order form.
\end{Lemma}

\begin{proof}
First proposition is easy to prove by contradiction. Second part, the candidate pair of polynomials only have one term of
single word-level output variable (say it is $R$) and this term is the last term under RATO, which means there is only one term with
$R$ in Spoly. Meanwhile in other polynomials from $J+J_0$ there is no such term containing $R$, so this term will be
kept to remainder $r$, in first degree.
\end{proof}

\begin{Example}
The elimination ideal for 5-bit SMPO (Ex.\ref{ex:SMPO}) could be rewritten under RATO:
\begin{align}
&(R_0,R_1,R_2,R_3,R_4)>(r_0,r_1,r_2,r_3,r_4)\nonumber\\&>(c_1,c_2,c_3,c_4,b_0,b_1,b_2,b_3,b_4)\nonumber\\
&>(a_0,a_1,a_2,a_3,a_4)>R>r>(A,B)\nonumber
\end{align}
Thus the candidate pair is
$(f_w,f_g), f_w = R_0+r_4+b_0\cdot a_1, f_g =R_0\alpha^5+R_1\alpha^{10}+R_2\alpha^{20}+R_3\alpha^9+R_4\alpha^{18} + R$.
Result after reduction is:
\begin{align}
&Spoly(f_w,f_g) \xrightarrow{J+J_0}_{+}\nonumber\\
&r_1+(\alpha)r_2+(\alpha^4+\alpha^2)r_3+(\alpha^3+\alpha^2)r_4+(\alpha^3)b_1a_1+(\alpha^4+\alpha^2)b_1a_2\nonumber\\
+&(\alpha^3+\alpha+1)b_1a_3+(\alpha^3+\alpha)b_1a_4+(\alpha+1)b_1A+(\alpha^4+\alpha^2+\alpha)b_2a_1\nonumber\\
+&(\alpha^4+\alpha^3+\alpha^2+\alpha)b_2a_4+(\alpha^3+\alpha^2+1)b_3a_1+(\alpha)b_3a_3\nonumber\\
+&(\alpha^2+\alpha+1)b_4a_1+(\alpha+1)b_4a_2+(\alpha^4+\alpha^2)b_4a_3+(\alpha^4+\alpha^3+\alpha+1)b_4a_4\nonumber\\
+&(\alpha^3+1)b_4A+(\alpha^4+\alpha^3+\alpha^2+1)a_1B+(\alpha^4+\alpha^3+\alpha^2+1)R\nonumber
\end{align}
The remainder satisfies Lemma \ref{lem:1}.
\end{Example}

\subsection{Bit-level Variable Substitution (BLVS)}
The remainder from \emph{Spoly} contains some bit-level variables, and our objective is to get a polynomial contains only word-level variables
(such as $R+\mathcal{F}(A,B)$). One possible method is to rewrite bit-level variables in term of function on word-level
variables, i.e. $a_i = \mathcal{G}(A)$, then do substitution. A Gaussian-elimination-fashion approach could be applied to
compute corresponding $\mathcal{G}(A)$ efficiently.

\begin{Example}
{\bf Objective}:\ Abstract polynomial $a_i + \mathcal{G}_i(A)$ from $f_0: a_0\alpha^5+a_1\alpha^{10}+a_2\alpha^{20}+a_3\alpha^9+a_4\alpha^{18}+A$.

First, compute $f_0^2: a_0\alpha^{10}+a_1\alpha^{20}+a_2\alpha^{9}+a_3\alpha^{18}+a_4\alpha^{5}+A^2$. Apparently variable $a_0$ can be
eliminated by operation 
\begin{align}
f_1 =& f_0\times \alpha^5 + f_0^2: \nonumber\\
&a_1+(\alpha)a_2+(\alpha^4+\alpha^2)a_3+(\alpha^3+\alpha^2)a_4\nonumber\\
+&(\alpha^4+\alpha^3+\alpha^2+1)A^2+(\alpha^2+\alpha)A\nonumber
\end{align}
Recursively eliminate $a_1$ from $f_1$, $a_2$ from $f_2$, etc. The final polynomial $f_4$ has the form 
$a_4 + \mathcal{G}_4(A) = a_4+(\alpha^4+\alpha^3+\alpha)A^{16}+(\alpha^3+\alpha^2)A^8+(\alpha^4+1)A^4+(\alpha^2+1)A^2+(\alpha+1)A$. Recursively substitute
$\mathcal{G}_4(A)$ back to $f_3$, etc. The result is a set of polynomials
\begin{displaymath}
\{g_i\ | \ g_i: a_i + \mathcal{G}_i(A)\}
\end{displaymath}
\end{Example}
In this approach it is easy to get word-level variable representation for each bit-level primary inputs. By substitution, a new polynomial in the form $R+\mathcal{F}(A,B)$
could be attained.

\textit{Discussion}\ \ (a) This approach won't conduct to a result reducing to 0, because \emph{Spoly}'s remainder contains information from the whole system,
the substitution will result something new, i.e. an abstraction of the system; (b) The Gaussian-like approach is a pre-processing of a variable, and could be immediately reused
to other words with the same length; (c) For $n$-bits SMPO, the fast GB is very efficient: for each cycle, first compute a \emph{Spoly} in $O(1)$, then do multi-division by $O(n)$ polynomials
(unknown complexity, maybe lower if using F4-style reduction?), the substitution cost at most $O(n^3)$ time (assuming $O(1)$ for each term). In total repeat for $n$ cycles.
The complexity should be much lower compared to Buchberger's algorithm.
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
%\bibliographystyle{abbrv}
%\bibliography{seq_verif}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
\appendix
\section{Normal Basis Multiplication}
\subsection{$\lambda$-Matrix}
$\lambda$-Matrix is defined with cross-product terms from multiplication. That is 
\begin{displaymath}
Product\ C = (\sum_{i=0}^{n-1}a_i\beta^{2^i})(\sum_{j=0}^{n-1}b_j\beta^{2^j}) = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}a_ib_j\beta^{2^i}\beta^{2^j}
\end{displaymath}
The expressions $\beta^{2^i}\beta^{2^j}$ are referred to as cross-product terms, and can be represented by
normal basis, i.e.
\begin{displaymath}
\beta^{2^i}\beta^{2^j} = \sum_{k=0}^{n-1}\lambda_{ij}^{(k)}\beta^{2^k}, \ \ \lambda_{ij}^{(k)} \in F_2.
\end{displaymath}
Substitution yields, get the expression for $k$-th digit of product:
\begin{displaymath}
c_k = \sum_{i=0}^{n-1}\sum_{j=0}^{n-1}\lambda_{ij}^{(k)}a_ib_j
\end{displaymath}
$\lambda_{ij}^{(k)}$ is the entry with coordinate $(i,j)$ in $k$-th $\lambda$-Matrix.

\begin{Example}
$\lambda$-Matrix: A binary $n\times n$ matrix $M$ could be employed to describe the "function"
 mentioned in Ex.\ref{ex:nb_multi}: $c_{n-1} = f(A, B) = A \cdot M \cdot B^T$, $B^T$ denotes vector transposition. 
More specifically, denote the matrix by \emph{$k$-th $\lambda$-Matrix}: $c_k = A \cdot M^{(k)} \cdot B^T$.
Then $c_{k-1} = A \cdot M^{(k-1)} \cdot B^T = rotate(A) \cdot M^{(k)} \cdot rotate(B)^T$, which means $M^{(k)}$ is generated
by right and down cyclic shifting $M^{(k-1)}$.

In $\mathbb{F}_{2^3}$ constructed by $\alpha^3 + \alpha + 1$, let $\beta = \alpha^3$, $N = \{ \beta, \beta^2, \beta^4\}$ 
is a normal basis. $0$-th $\lambda$-Matrix
\begin{displaymath}
M^{(0)} = \left(
\begin{array} {lcr}
0 & 1 & 0\\
1 & 0 & 1\\
0 & 1 & 1
\end{array} \right).
\end{displaymath}
i.e.,
\begin{displaymath}
c_0 = (a_0\  a_1\  a_2)\left(
\begin{array} {lcr}
0 & 1 & 0\\
1 & 0 & 1\\
0 & 1 & 1
\end{array} \right)\left(
\begin{array} {lcr}
b_0\\
b_1\\
b_2
\end{array} \right).
\end{displaymath}
\end{Example}
\subsection{Optimal Normal Basis}
\begin{Definition}
The number of non-zero entries in $\lambda$-Matrix is known as  \emph{Complexity$(C_N)$}.
\end{Definition}
\begin{Theorem}
If N is a normal basis for $\mathbb{F}_{p^n}$ with $\lambda$-matrix $M^{(k)}$, then non-zero entries in 
matrix $C_N\geq 2n-1$.
\end{Theorem}
Proof omitted.
\begin{Definition}
If there exists a set of normal basis satisfying $C_N = 2n - 1$, this normal basis is named as
\emph{Optimal Normal Basis (ONB)}.
\end{Definition}
There are 2 types of optimal normal basis existing.
Rules for creating Type-I ONB over $\mathbb{F}_{2^n}$ are:
\begin{itemize}
\item n+1 must be prime.
\item 2 must be primitive in $\mathbb{Z}_{n+1}$.
\end{itemize}
Rules for creating Type-II ONB over $\mathbb{F}_{2^n}$ are:
\begin{itemize}
\item 2n+1 must be prime. And either
\item 2 must be primitive in $\mathbb{Z}_{2n+1}$, OR
\item $2n+1 \equiv 3 \ \ mod\ \  4$ AND 2 generates the quadratic residues in $\mathbb{Z}_{2n+1}$
\end{itemize}
There are corresponding criteria for simply creating $\lambda$-Matrix for either type of ONB:
\begin{Lemma}
Type-I ONB implies a $\lambda$-Matrix that each nonzero entry $M_{i,j}$ satisfies
\begin{numcases}{}
2^i + 2^j = 1 \bmod  n+1\notag\\
2^i + 2^j = 0 \bmod n+1\notag
\end{numcases}
Type-II ONB implies a $\lambda$-Matrix that each nonzero entry $M_{i,j}$ satisfies
\begin{numcases}{}
2^i + 2^j = 1\bmod\  2n+1\notag\\
2^i + 2^j = -1 \bmod  2n+1\notag\\
2^i - 2^j = 1 \bmod  2n+1\notag\\
2^i - 2^j = -1 \bmod 2n+1\notag
\end{numcases}
\end{Lemma}
Proof omitted. (Issues here: We knew type-I \& II definition can deduce corresponding $\lambda$-Matrix, how about
reverse implication? i.e. can we prove the equivalence?)

\subsection{Design a Normal Basis Multiplier using $\lambda$-Matrix}
Imitating the structure of SMPO, just specify the gates' connections for the first clock cycle, then following cycles are
automatically completed by cyclic shifting operands $A$ and $B$. The whole design procedure is based on a $n \times n$ $k$-th $\lambda$-Matrix.

First figure out the first row (\emph{row} 0) in $\lambda$-Matrix, for any nonzero entry $M_{0,j}$ (there will be only 1 if taking $0$-th $\lambda$-Matrix), place an AND gate
$a_j\land b_0$. Connect different $a_j$ with a XOR gate before place the AND gate if there are 2 or more nonzero entries in this row;

Secondly, repeat this for each row. Note for nonzero entry $M_{i,j}$, corresponding index of $a_u$ and $b_v$ is incremented by $i$, i.e. $u = j + i \pmod n$, $v = i + i \pmod n$;

Finally, connect the output of AND gate for row $i$ to one input of a XOR gate, then connect the output of XOR gate to a register/flip-flop. The output of register/flip-flop is connected
to the other input of XOR gate, but at next row. All these gates and connections for row $i$ is called unit $R_i$. (Fig.\ref{fig:SMPO})

\section{Find Optimal Normal Basis}
\subsection{Transforming to SAT}
From last section we learned it is easy to compute $\lambda$-Matrix when given certain type of optimal normal basis. However in our approach, the mapping between 
optimal normal basis and standard basis should be explicitly specified. In other words, let \emph{Optimal Normal Element} $\beta = \alpha^k$, the exponential $k$ on
primitive element $\alpha$ needs to be calculated. Otherwise word-level variable interpreting polynomial $a_0\beta + a_1\beta^2 + \cdots + a_{n-1}\beta^{2^{n-1}}$
will be an unknown polynomial and GB computation is impossible.

Currently finding the optimal normal element $\beta$ in $2^n$ (over field $\mathbb{F}_{2^n}$) elements is NP-hard. However, it is possible to improve the performance 
by transforming this problem to SAT problem which can be solved by efficient SAT solver. This is what our tool is managing to do.

Given 0-th $\lambda$-Matrix $M^{(0)}$, other $\lambda$-Matrices can be obtained by cyclic right-down shifting, i.e. entry $M_{i,j}^{(k)} = M_{i-k,j-k}^{(0)}$. Note all
index operation results should modulus $n$. $\beta^3$ is an element which can be represented as product of normal basis multiplication: $\beta^3 = \beta\cdot\beta^2$.
Consider the following equation
\begin{displaymath}
c_k = (1\  0\ 0\ \cdots\ 0)\  M^{(k)} \left(
\begin{array} {lcr}
0\\
1\\
0\\
\vdots\\
0
\end{array} \right)
\end{displaymath}
$c_k$ is the $k$-th coefficient in normal basis expression of $\beta^3$: $\beta^3 = c_0\beta + c_1\beta^2 + \cdots + c_{n-1}\beta^{2^{n-1}}$.
Thus, $\beta^3$ could be written as
\begin{displaymath}
\beta^3 = \sum_{k=0}^{n-1} M_{0,1}^{(k)}\beta^{2^k} = \sum_{k=0}^{n-1} M_{-k,1-k}^{(0)}\beta^{2^k}
\end{displaymath}
Meanwhile, the standard basis representation of optimal normal element can be assumed as
\begin{displaymath}
\beta = a_0 + a_1\alpha + \cdots + a_{n-1}\alpha^{n-1}
\end{displaymath}
Do a substitution, vanish every coefficient. Finally we can get a linear system of $n$ boolean equations (composed by XOR and AND), with $n$ unknowns $a_0,a_1,\dots,a_{n-1}$. 
Examples are in the write-up by Florian's student.

\subsection{Solving XOR-SAT}
Solving the linear system of boolean equations is a tricky problem, because ordinary SAT solver only support Conjunctive Normal Form (CNF) inputs; while a XOR-SAT based solver 
\emph{CryptoMiniSAT} only support clauses composed by single literals and XOR connectives ($a\oplus b \oplus (c\land d)$ is illegal input), I call it "XOR-CNF". Basically there are 2 approaches 
to think about our tool flow:

(a) Use another engine (Commercial Symbolic Computing System such as \emph{Wolfram Mathematica}, or Circuit Based Synthesizer such as \emph{ABC}) to transform all equations to CNF clauses.
Advantage: Convenient, even no need to write our own tool in scripts; Disadvantage: cannot guarantee the efficiency, CNF may explode if we have hundreds of equations with hundreds of 
XOR terms.

(b) Imitate clause learning technique in CDCL, create new clauses to satisfy the input format of \emph{CryptoMiniSAT}. Advantage: \emph{CryptoMiniSAT} is a tool with performance enhancing
on XOR clauses, and since the number of XOR clauses will not explode (I will explain later), maybe the efficiency can be guaranteed (somehow); Disadvantage: need to build our own tool for experiments.

Let me have a tiny statement on the second choice. First is the approach, maybe better illustrated by an example:
\begin{Example}
Change boolean equation $a\oplus b\oplus c \oplus bc = 0$ to XOR-CNF. First, a SAT-solver always want whole CNF to be TRUE. Fortunately, we can negate a XOR clause by just negate one literal:
$\overline{a}\oplus b\oplus c \oplus bc = 1$. Then we need to solve the non-single-literal term. Replace term $bc$ with new literal $d$, and create a new clause to make $d=bc$, which is
$(\overline{b}\lor\overline{c}\lor d)\land(b\lor\overline{d})\land(c\lor\overline{d})$, adding 3 short clauses.
\end{Example}
You may ask why the number and size of clauses won't explode, the answer is easy: we only have 2-literal terms in worst case! Note all equations come from 
$$\beta^3 + c_0\beta + c_1\beta + \cdots + c_{n-1}\beta^{2^{n-1}} = 0$$
Even exponentials will never generate higher order coefficients in $\mathbb{F}_{2^n}$, so the only term generates higher-order coefficients is
$$\beta^3 = (a_0+a_1\alpha + \cdots a_{n-1}\alpha^{n-1})(a_0+a_1\alpha^2+\cdots + a_{n-1}\alpha^{2(n-1)})$$
Apparently cross-product terms here have at most 2 literals (degree 2).

Acknowledging this fact, for $n$ unknowns there will be at most $\dbinom{n}{2} = O(n^2)$ 2-literal terms, so the number of newly created clause will also be $O(n^2)$, with fixed tiny size.

By either approach, we finally get one legal assignment for bit vector $(a_0,a_1,\dots,c_{n-1})$ from the SAT-solver, by that we can compute $\beta$. We can directly use it as ONB, or setup a loop to keep squaring 
$\beta$ to find the minimum $k$ for $\alpha^k$, and adopt that one as ONB.

\textit{Discussion}\ \ The tool flow is feasible enough, so only issue is the soundness from the ONB theory to linear system of boolean equations. I asked following questions to Florian's student:
1. Solutions for linear system of equations in $\mathbb{F}_2$ (or say $\mathbb{Z}_2$). For example, in last page of ONB.pdf194 KB, there are 5 variables and 5 equations, provide 5 conjugate solutions forming a set of (optimal) normal basis.
Is it possible to dig into this, say to prove all solutions are conjugates $(x, x^2, x^4, x^8, \dots)$?
2. You chose $\beta^3$ to deduce those equations. Does solution to $\beta^3$ necessarily cover (or equal to) solutions for given lambda-Matrix M0? In other words, if I choose $\beta^5$ or $\beta^6$ to construct those equations, can I get the same answer?

If above concerns are resolved, we can come up with a complete proof.
\balancecolumns
% That's all folks!
\end{document}