### 2014-04-25 sequential/latex/DAC2013/xiaojun.tex

\section{Further discussion on Computer Algebra}
\label{sec:moretheory}

Bit level circuit variables including primary inputs/outputs can only take value
within $\Z_2$. They can also be written in functions of other variables and constants.
In computer algebra, circuit variable $a$ can be described as a polynomial, that is
$a = f(x_0, x_1, \dots, x_{n-1}) \in {\mathbb{F}}[x]$. Moreover, circuit variable can
have multiple specifications, they can be written in polynomials and included in an
ideal $I = \{f_0, f_1, \dots, f_k\}$. The varieties of this ideal are possible 
assignments of circuit variables. Similarly for $n$ bits world-level circuit 
variables, the varieties are elements from $\Fkk$. Thus it's convenient to use
ideals and varieties to describle circuit variables. \\

\begin{Definition}
({\bf Sum of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the 
{\bf sum} of $I$ and $J$, denoted by $I + J$, is the set
  \begin{equation}
  I + J = \{f + g : f \in I \ and\  g \in J\}.
  \end{equation}
Furthermore, if $I = \langle f_1, \dots, f_r\rangle$ and 
$J = \langle g_1, \dots, g_s\rangle$, then 
$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$.
\end{Definition}

\begin{Definition}
({\bf Product of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then the
{\bf product} of $I$ and $J$, denoted by $I \cdot J$, is defined to be the ideal generated 
by all polynomials $f \cdot g$ where $f \in I$ and $g \in J$. Furthermore, let
$I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$, then
  \begin{equation}
  I \cdot J = \langle f_ig_j : 1 \leq i \leq r, 1 \leq j \leq s\rangle .
  \end{equation}
\end{Definition}

\begin{Definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then $I:J$
is the set
  \begin{equation}
  \{f \in k[x_1, \dots, x_n] : fg \in I, \forall g \in J\}
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{Definition}

According to definitions above, it is possible to calculate ideal sum, product and 
quotient based on ideal generators. Additional conclusion about union, 
intersection and complement of varieties can be attained below: \\

\begin{Theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $k[x_1, \dots, x_n]$, then ${\bf V}(I + J) = {\bf V}(I)
\bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) = {\bf V}(I) \bigcup {\bf V}(J)$.
\end{Theorem}

\begin{Theorem}
\label{thm:quotient}
Let $V$ and $W$ be varieties in $\Fkk$. Then ${\bf I}(V) : {\bf I}(W) = 
{\bf I}(V - W)$.
\end{Theorem}

For \ref{thm:quotient}, let $V = \Fkk = {\bf V}(\langle vanishing \  polynomials
\rangle)$, then $V - W$ is the \textbf{complementary set} of variety $W$.\\

Based on these conclusions it's easy to get a circuit variable's specifications
or possible assignments from the other known information.

\section{Sequential circuits and Finite State Machine}
\label{sec:fsm}
\subsection{Introduction to Finite State Machine:}
Borrow from somewhere.
\subsection{Symbolic Image Computation:}
Borrow from somewhere.
\subsection{Implicit State Enumeration:}
BFS algorithm below:
\begin{algorithm}[hbt]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}

  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cdot \overline{reached}$\;
  	$reached \gets reached + new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}\label{alg:gb}
\end{algorithm}

\section{Experiments using new approach}
\label{sec:exp}

First example is reachable state enumeration of a 2-bit finite state machine.
Breadth-First-Search traversal (mentioned in \ref{sec:fsm}) is adopted.

\begin{Example}
\label{ex:traversal}
Sample circuit is described below with gate-level implement and state transition
graph.
  \begin{figure}[hbt]
    \centerline{\includegraphics[scale=0.3]{./fsm_ckt.png}}
    \caption{Sample circuit}
  \label{fig:fsmckt}
  \end{figure}

  \begin{figure}[hbt]
    \centerline{\includegraphics[scale=0.3]{./fsm_stg.png}}
    \caption{State Transition Graph}
  \label{fig:fsmstg}
  \end{figure}

The following mapping from Boolean operations to Galois field $\mathbb{F}_2$ 
operations is straightforward. Assuming $a$ and $b$ are both bit-level variables:
\begin{itemize}
  \item $\overline{a} \rightarrow 1 + a$
  \item $a\ and\ b \rightarrow ab$
  \item $a\ or\ b \rightarrow a + b + ab$
\end{itemize}
Comparing to the quantification approach in \ref{sec:fsm}, using Gr\"obner bases
approach from \ref{sec:theory}. Construct an elimination ideal with following 
polynomials: $f_1: 1 + XNOR(t_0, OR(p\cdot q, (1 + x_i)\cdot(1 + p)\cdot(1 + q))); 
f_2: 1 + XNOR(t1, OR(xi\cdot(1+p),p\cdot(1+q))); f_3: S - p - q\cdot\alpha; 
f_4: T - t_0 - t_1\cdot \alpha$ and vanishing polynomials $f_5: p^2 - p; 
f_6: q^2 - q; f_7: t_0^2 - t_0; f_8: t_1^2 - t_1; f_9: S^4 - S; f_{10}: x_i^2 - x_i; 
f_{11}: T^4 - T$, where $S$ and $T$ are 2-bit word representing current state
and next state as defined in $f_3$ and $f_4$, $f_1$ and $f_2$ describe the 
transition relations. At last the initial state should also be included in 
this elimination ideal. For example initial state is $S_3$, then last polynomial
is $f_{12}: S + 1 + \alpha$. According to elimination theorem \ref{thm:elim},
let the lex ordering be $\{all\ others\} > T$, calculate Gr\"obner basis, the 
result must contain a polynomial with only one variable $T$, which can indicate
the possible assignments of next state. \par

Apply this approach to calculate image in BFS traversal \ref{alg:BFS}.
In this example, $to^i$ and $reached$ are both ideals containing only one
generator, i.e. the polynomial with single variable $T$; universal set is
varieties from ideal $\langle T^4 - T\rangle$. Use \ref{thm:unionintersect} and
\ref{thm:quotient} to complete \ref{alg:BFS}, the final return value is 
polynomial $to^3 = T^4 + T$, which means all states are rechable from state $S_3$.
\end{Example}

The second example shows the application of the new approach on sequential 
arithematic circuits function verification.

\begin{Example}
\label{exp:multiplier}

The following design is Sequential Multiplier with Parallel Output (SMPO),
a Normal Basis multiplier based on Massey-Omura algorithm. Inputs and outputs
are all 5-bit word taking value from ${\mathbb F}_{2^5}$. After loading operands
 A and B, setting all output latches to 0 and running for 6 iterations, the 
output should be $R = A\cdot B$ (mod $x^5 + x^2 + 1$).
  \begin{figure}[hbt]
    \centerline{\includegraphics[scale=0.3]{./SMPO.png}}
    \caption{5-bit SMPO}
  \label{fig:SMPO}
  \end{figure}

Similarly, build elimination ideal for all gates/operations and induce
initial states of latches. However, instead of eliminating all variables
to one, this example adopts abstraction term order from \ref{thm:abs}
and keeps the polynomial contains function between output and inputs.
Here the lex ordering is $\{others\} > R > \{A, B\}$, and objective
polynimial is $R + \F(A, B)$. \par

Apply this approach to calculate image in BFS traversal, but modify 
\ref{alg:BFS} to make it adapt 6 steps reachable states enumeration.
The result is $R + AB$, which validates the function of this circuit.

\end{Example}