### 2014-04-25 sequential/latex/ICCAD2014/abstract_fmcad.tex

\begin{abstract}
Sequential circuits that implement Galois field arithmetic over
$\Fkk$ are prevalent in cryptography and error control coding, where
the datapath size $k$ is very large. Formal verification of sequential
arithmetic circuits with large datapath size is beyond the
capabilities of contemporary verification techniques. To address this
problem, this paper introduces a new concept of implicit state
enumeration of finite state machines (FSMs) performed at the
word-level. Using algebraic geometry, we show that the state-space of
a sequential circuit can be encoded, canonically, as solutions to a
multivariate word-level polynomial $\F(A)$ in $\Fkk$, where $A = {a_0,
\dots, a_{k-1}}$ represents the $k$-bit state-register (word). Subsequently,
the Groebner basis algorithm, along with elimination ideals and quotients 
of ideals can be employed for FSM traversal. We apply our approach 
to verify sequential Galois field multipliers with large datapath sizes, 
where contemporary techniques prove to be infeasible.
\end{abstract}