### 2014-04-25 sequential/latex/Word-Level RTL.tex

%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[conference,11pt]{IEEEtran}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

\usepackage{graphicx}

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\usepackage[tight,footnotesize]{subfigure}
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\usepackage[caption=false]{caption}
%\usepackage[font=footnotesize]{subfig}
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\usepackage[caption=false,font=footnotesize]{subfig}
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Word-Level RTL Equivalence Checking Using Word-Level Constraint-Models}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{Xiaojun Sun}
\IEEEauthorblockA{Electrical and Computer Engineering\\
University of Utah\\
Salt Lake City, Utah\\
Email: jason.sj.sun@gmail.com}
\and
\IEEEauthorblockN{Shaofan Qi}
\IEEEauthorblockA{Electrical and Computer Engineering\\
University of Utah\\
Salt Lake City, Utah\\
Email: chi.cedric@gmail.com}
\and
\IEEEauthorblockN{Yi Guo}
\IEEEauthorblockA{School of Computing\\
University of Utah\\
Salt Lake City, Utah\\
Email: broroguoyi@hotmail.com}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

%\begin{abstract}
%%\boldmath
%The abstract goes here.
%\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\begin{abstract}
%\boldmath
To check RTL equivalence using word-level constraints,\emph{ LPSAT} is introduced to solve the satisfiability (SAT) problem for complex RTL-level
designs, which contain word-level arithmetic operators and Boolean logics. Our project uses linear-program solvers for constraint satisfaction.
The word-level SAT solver, called LPSAT, based on linear programming. In this approach, both word-level arithmetic constraints and control logic constraints are put together into a common linear programming domain.
\end{abstract}

\section{Introduction}
% no \IEEEPARstart
The Boolean satisfiability problem (SAT) belongs to the class of NP-complete problems, with algorithmic solutions having exponential worst-case complexity. Word-Level RTL equivalence checking using linear constraint-models. Linear Programming (LP) is a special case of a constrained optimization problem in which both the objective function and constraints are linear equations. In addition, variables in LP are assumed to be non-negative.

In this project, our major target is to implement the equivalence checking with word-level RTL and linear constraint. Register-transfer level (RTL) is referred as a design abstraction which models a synchronous digital circuit in terms of the flow of digital signals (data) between hardware registers, and the logical operations performed on those signals. Word-level abstraction or constraint-models are usually used to check world-level RTL equivalence. This project aims at proving the equivalence between two given RTLs, using integer linear-programming solvers for constraint satisfaction.

Satisfiability of complex word-level formulas described at the RTL is usually solved in the Boolean domain, with Boolean solvers. However, these engines often show a poor performance for data path verification. Instead of solving the problem at the bit-level, a method is proposed to transform conjunctions of bit-vector equalities and inequalities into sets of integer linear arithmetic constraints. Integer linear constraint solvers are used as a decision procedure for bit-vector arithmetic. Linear Programming (LP) is a special case of a constrained optimization problem in which both the objective function and constraints are linear equations. In addition, variables in LP are assumed to be non-negative.

 Bit-vectors are used to represent variables and their values by RTL-HDLs.  Important arithmetic bit-vector operators are for instance: \emph {negation, addition, multiplication, extension, concatenation, and shifting}. These operators are often used to model data paths of hardware designs.  Many design automation tools operating on HDLs, including logic verification, test generation, and synthesis, have to solve the satisfiability problem for word-level formulas.  Solving the SAT problem directly on the word-level, as opposed to ﬁrst translating it into the Boolean domain, bears a high potential for optimization for those tools.  In the RTL, we should build the linear model of the operators shown in table 1.

In this project, we first find all these operators’ linear constraint model, and the way to linearize arithmetic operators and Boolean gates. Then build a program to cover the circuit design, which is RTL-level (verilog file). Then use the Linear Programming (LP) technique (AMPL, IBM CPLEX) to model and solve this SAT problem.  ILOG CPLEX is a tool for solving linear optimization problems, commonly referred to as Linear Programming (LP) problems.[1]

\begin{table}[!t]
% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Operators}
\label{work}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{|c|c|c|}
\hline
Arithmetic Operators& Bitwise Operators & Logical Operators \\
\hline

$+, -, *,  \% $ &   \& , $\vert$ , \^{} , \~{} , \^{}\~{} , \~{}\^{}   & ! , \&\& , $\vert$$\vert$  \\
\hline
\hline
Relational Operators& Equality Operators & Shift Operators \\
\hline
$>, < , >= , <= $ &   $ == , != $   & $ >> , << $  \\
\hline
\hline
Conditional Operators&& \\
\hline
$?: $ &&\\
\hline
\end{tabular}
\end{table}
%\hfill mds
%\hfill January 11, 2007

\section{Linear Programming}
Linear Programming (\emph{LP} or \emph{linear optimization}) is a mathematical method for determining a way to achieve the best outcome in which both the objective function and constraints are linear equations.

Linear programs are problems that can be expressed in canonical form:

\begin{eqnarray*}
\emph{Minimize} & : & c^T X\\
\emph{Subject  to} & : & AX <= b\\
 & & X>=0\\
\end{eqnarray*}
% this part is trivial
where $X$ represents the vector of variables (to be determined), $c$ and $b$ are vectors of (known) coefficients, $A$ is a (known) matrix of coefficients, and is the matrix transpose. The expression to be maximized or minimized is called the objective function ($c^T X$ in this case). The inequalities $Ax <= b$ are the constraints which specify a convex polytopes over which the objective function is to be optimized. In this context, two vectors are comparable when they have the same dimensions. If every entry in the first is less-than or equal-to the corresponding entry in the second then we can say the first vector is less-than or equal-to the second vector.[2]

\section{Modeling}
This part is using Linear Programming (LP) technique to model the SAT problem for a complex design composed of the arithmetic and Boolean logic blocks. Basic arithmetic word-level operators include: adders, subtractors, comparators, , shifters, and multipliers.  Any other word-level operators can be represented in
terms of these operators using Boolean connectors (AND, OR, NOT).

	\subsection{ Adder/Subtractor}
	Addition and subtraction are both linear operators and can be trivially represented as equality constraints:
	\begin{eqnarray*}
	& &C=A \pm B \Rightarrow C - A  \mp B = 0
	\end{eqnarray*}

	\subsection{ Comparison operators}
	There are six types of comparison operators:$ \leq , < , \geq, > , = , \neq$. For a comparator, $A = (B == C) \emph{ or } A = (B < C) \emph{ or }  A = (B > C)$, variables
$x , y$ are introduced to indicate that $(B==C) \Leftrightarrow (x y = 00)$ , $(B>C) \Leftrightarrow (x y = 10)$ and $(B<C) \Leftrightarrow (x y = 01)$
$$\left\{
\begin{array}{lcl}
{A = 1- x -y} \\
{B - C+ L \cdot (1-x) \geq 1} \\
 {B-C+L\cdot x \leq 0}\\
 {C -B +L\cdot (1-y) \geq 1}\\
  {C -B +L\cdot y \leq 0}\\
  {x +y \leq 1}
\end{array} \right. $$
	Here $L$ is a constant and satisfying the constraint $L \geq max(B,C)$. We assume the $max(B,C)$ is used to denote a constant,  it means the maximum value of the upper bound of A or B $(\emph {in our project, we made } L = 2^n -1)$.

	\subsection{Multiplier}
	Multiplication is a non-linear operator, like $A = B \times C$. So one of $B$ or $C$ must be expanded into binary variables. The best candidate for the expansion is the one which is driven
by the Boolean part. Let \emph{B} be an n-bit variable to be expanded, and let $P_i = B_i \times C$ be a partial product, and $B_i$ has to be declared as a binary variable. And the $A = \sum_{i=0} ^{n-1} 2^i \cdot P_i$, and for each $i \in [0,1,\ldots,n-1] $,
$$\left\{ \begin{array}{l}
{P_i -L \cdot B_i \leq 0} \\
{P_i - C+ L \cdot (1-B_i) \geq 0} \\
{0 \leq P_i \leq C}
\end{array} \right. $$
[3]

\section{Implementations}
For our implementation, we chose \emph{Perl} to implement converter, which will translate the RTL-level verilog to our constraints file.
The algorithm is: First, open the RTL verilog file as input, read in one line, find if it's an assertion, record signal name and its width;
Second, go to signal assignment part (usually right of first``=''), split whole sentence into words by space(s);
Then, record the number of operators as counter for iteration, each loop find a arithmetic unit where left operand does not contain ``)'' and right operand does not contain ``('',
calculate this unit with given operands and operator and their bits width, print out the constrains into result file;
Finally, judge if we need to generate a temporary variable to record the result, then replace the original expression ``(A op B)'' with new variable C in the string, continue iteration until all operations are done.
This is a \emph {pseudo-code}:

\begin{verbatim}
$readin = <>; 
\end{verbatim}
 \emph {read in a line}
\begin{verbatim}
@items = split /\s+/, $readin;  
\end{verbatim} 
\emph {split with spaces to get words}
\begin{verbatim}
if (Not an assertion) {next;}  
\end{verbatim} 
\emph {an assertion requires ``=" behind signal name}
\begin{verbatim}
$def_var = $items [1]; 
$wid_signal = /\[(\d+)\:/; \end{verbatim}  \emph {extract width of signal} \begin{verbatim} 
$var_list {$def_var} = $wid_signal;  \end{verbatim}  \emph {store this info into a hash table} \begin{verbatim} 
While (find basic operators)
{
    Operator count ++;
\end{verbatim}
\emph{find out the smallest unit, which just contains one operator}
\begin{verbatim}
	if (judge unit == FALSE) {next;}
\end{verbatim}
	\emph{Write linear constrains to output file}
\begin{verbatim}
    &extract($picked string) 
  $op_left, $op_right, $wid_left, $wid_right;  \end{verbatim}  \emph {extract 2 operands and their width (bits No.)} \begin{verbatim} 
if ($op_left || $op_right contain [i:j] or -) 
  {&unary_op($oprand);}  \end{verbatim}  \emph { judge if we need to process unary operand first } \begin{verbatim}
if (exsit $var_list {C} == FALSE) 
  {$obj = "my_var_$tmp_count";} 
&write_constrains 
  ($op_left, $op_right, $wid_left, 
	$wid_right, $obj) 
\end{verbatim}
\emph{remove excess brackets, store into new element. And replace the unit with an temporary variable}
\begin{verbatim}
    Find immediate (){},Remove(){};  \end{verbatim}  \emph {if operand is ``," for catenation then apply "\{\}" } \begin{verbatim} 
	replace substr $items[i-1][i][i+1] with $obj;  \end{verbatim}  \emph { replace for binary operation } \begin{verbatim}
}
\end{verbatim}

\section{Experimentation and Results}
The experimentation of our prototype required a little revise of the output constraints manually. Also the constraint of multiplier is using some logical function such as \emph{floor()}. The way we try to linearize these logical functions by introducing new temporary \textbf {Boolean} variables to regulate constraints. We conducted two tests (for translating a same verilog file but using different solvers). The first one with of standard AMPL/CPLEX:
\begin{verbatim}
ampl: model P0.mod;
ampl: option solver cplex;
ampl: solve;
CPLEX 11.2.0:
Objective = find a feasible point.
\end{verbatim}

 Second test was using MINOS, which is AMPL default solver :
\begin{verbatim}
ampl: option solver minos;
ampl: solve;
MINOS 5.51:
6 iterations
Nonlin evals: constrs = 6, Jac = 6.
Objective = find a feasible point.
\end{verbatim}

\section{Conclusion}
This project offered a successful demonstration of Word-Level RTL Equivalence Checking using Perl and AMPL/CPLEX. It red verilog file and converted the RTL-level descriptions into Linear Programming constraints. It also successfully verified the assignment with  AMPL/CPLEX.

\section{Future work}
While our prototype offers a working solution, it is far from its full potential. In the future work, we can add more LPATPG function into our project. We can also do more tests on boolean logical test. Designing a fully automatic converter is also in our a further plans. And build a user friendly GUI can be scheduled at same time.

\begin{thebibliography}{1}

\bibitem{ref1}
Z. Zeng, P. Kalla, M. J. Ciesielski, \emph{``LPSAT: A Uniﬁed Approach to RTL Satisﬁability,” } ICCAD'08, 2008.

\bibitem{ref2}
\emph{Linear Programming} http://en.wikipedia.org/wiki/Linear\_programming

\bibitem{ref3}
Z. Zeng \emph{``FUNCTIONAL TEST GENERATION BASED ON WORD-LEVEL SATISFIABILITY ,” } Dissertation of Zhihong Zeng 2002.

\end{thebibliography}

% that's all folks
\end{document}

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex,
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

%\section{Conclusion}
%The conclusion goes here.

% conference papers do not normally have an appendix

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

% that's all folks
\end{document}