### 2014-06-06 today/conjecture.tex
modify bugs in 2bit traversal and add conjecture tex

\documentclass[12pt,letterpaper]{article}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd}
\usepackage{fullpage}
\usepackage{fancyhdr}
\usepackage{cases}
\usepackage{geometry}
\geometry{left=2.5cm,right=2.5cm,top=1cm,bottom=2cm}
\pagestyle{fancyplain}
\headheight 15pt
%\lhead{Xiaojun Sun\ (u0741290)}
%\chead{\textbf{\Large Homework \asgnname}}
%\rhead{\today}
%\headsep 10pt

\begin{document}

\title{Conjecture on multivariate generator ideals' quotient}
\author{Xiaojun Sun}
\maketitle

\section{Problem Description}
From Prof. Enescu's last write-up, our conjecture on univariate generator ideals' quotient has been
proved. Let $K = \mathbb F_q$, where $q=p^e$, $p$ is a prime number. We have:\par
\indent \\
\textbf{Lemma}\ \ Let $f,g\in K[x]$, then $(f:g) = (\frac{f}{gcd(f,g)})$.\par
\indent \\
\textbf{Proposition}\ \ Let $J$ be an ideal of $K[x]$ and $J_0 = (x^q - x)$. Then
$$V_K(J_0:J) = V_K(J_0)\setminus V_K(J).$$
\indent \par
Note above assertions dealt with ideals $J_0, J$ with only 1 generator and this generator is limited
to univariate polynomial about $x$. 
Now we would like to extend this conjecture on ideals with multivariate polynomial generators. Of course
those generators must be written in specific form. Here we give an example, which covers
typical requirement for our verification theory.\par
\indent\\
\textbf{Conjecture}\ \  Assume we have 3 ring variables $x,y,z$, and an ideal $J$ composed by 3 generators:
$J = \langle f, v_x, v_y\rangle$. Here $f$ has the form $f= z+\mathcal F(x,y)$, where $\mathcal F$
is a 2-argument function, for example $f = z+(x^2y+xy^2)$. And $v_x, v_y$ are "vanishing polynomials" about
$x, y$: $v_x = x^q - x, v_y = y^q - y$. Now we have another generator serving as "universal set"
$J_0 = \langle v_z, v_x, v_y\rangle$, $v_z = z^q - z$. (If necessary, $J$ can also include $v_z$ as 4th generator)
We conjecture that
$$V_K(J_0:J) = V_K(J_0)\setminus V_K(J).$$
And we believe this conjecture is scalable such that it is also valid for situation 
$J = \langle f, v_x,v_y,v_u,v_w\rangle$, $f = z + \mathcal F (x,y,u,w)$.
\end{document}