### 2015-05-01 techwriting/ASPDAC/abs.tex
all write-up from virtue machine

{\it Abstract -- Sequential Galois field ($\Fkk$) arithmetic circuits
take $k$-bit inputs and produce a $k$-bit result, after $k$-clock
cycles of operation. Formal verification of sequential
arithmetic circuits with large datapath size is beyond the
capabilities of contemporary verification techniques. To address this
problem, this paper describes a verification method based on algebraic
geometry that: i) implicitly unrolls the sequential arithmetic circuit
by multiple ($k$) clock-cycles and ii) represents the function
computed by the state-registers of the circuit, canonically, as a
multi-variate word-level polynomial in $\Fkk$. Our approach employs
the \Grobner basis theory with a specific elimination ideal. Moreover,
an efficient implementation is described to identify the $k$-cycle
computation performed  by the circuit at word-level. We demonstrate
the feasibility of our approach by verifying up to 100-bit sequential
Galois field multipliers, whereas conventional techniques fail beyond
23-bit circuits.  }