### 2015-05-01 techwriting/ASPDAC/intro_aspdac.tex

\section{Introduction}
\label{sec:intro}

Galois field (GF) arithmetic finds application in many areas such as 
cryptography, error control coding, VLSI testing, etc. In most
hardware applications, fields of the type $\Fkk$ are widely
chosen. Such {\it binary} GFs are $k$-dimensional extensions
of the base field ${\mathbb{F}}_2$; this allows for the design of
efficient (AND-XOR) arithmetic architectures and algorithms for
hardware design. For example, in elliptic curve cryptography, the
operations of encryption, decryption and authentication rely on
polynomial computations over $\Fkk$. In many applications, however,
the datapath size (bit-vector operand size) $k$ can be very
large. Therefore, for efficient implementation, many custom VLSI
architectures for polynomial GF aritmetic have been devised
\cite{wu:2002} \cite{acar:1998} \cite{agnew1991implementation}
\cite{RHmulti}. Due to their deployment in communications and
security-related applications, there is a critical need to {\it formally
verify} the correctness of hardware implementations of GF arithmetic.

Formal verification of such specialized and custom GF arithmetic
architectures remains a challenge. While there have been recent
advancements in formal verification of {\it combinational} GF
arithmetic circuits \cite{ibm:blueveri} \cite{lv:tcad2013}
\cite{pruss:dac14}, formal verification of sequential GF circuits has
not been addressed.

The high complexity

 This introduce many
design and verification challenges