### 2016-05-12 unsat/paper/abstract.tex

\begin{abstract}
Finding small unsatisfiable subformulas (unsat cores) of infeasible
constraint problems is an active area of research. The problem has
been widely investigated in the propositional SAT domain. Analogous
investigations in the polynomial algebra domain are, however,
somewhat lacking. This paper investigates an algorithmic approach to
identify a small unsatisfiable core of an ideal in polynomial rings
with coefficients from any field, where the ideal is found to have an
empty variety. We show that such a core can be identified by means the
Buchberger's algorithm and its variations. We identify certain
conditions that are helpful in ascertaining whether or not a
polynomial from the given generating set is a part of the unsat
core. Our algorithm cannot guarantee a minimal unsat core; hence the
paper discusses opportunities for refinement of the identified
core. The approach is demonstrated on a few examples, and experiments
are conducted using an implementation of our algorithm within the
SINGULAR computer-algebra tool. 
\end{abstract}