### 2016-10-19 dissertation/recha.tex
dissertation:part reacha:update 1

\chapter{Word-level Traversal of Finite State Machines using Algebraic Geometry}
\label{ch:reacha}
Reachability analysis is a basic component of sequential circuit verification, esp. for 
formal equivalence checking and model checking techniques. Concretely, in modern synthesis 
tools, in order to improve various performance indicators such as latency, clock skew or power
density, some major refactoring and attachments are made on original designs. 
Those modifications may introduce malfunctions or glitches to the whole logic. 
In localized simulation or formal verification (esp. equivalence checking), the modifications may be 
denied since the malfunctions or glitches are considered as ``faults" in this circuit.
However, if the circuit behavior is carefully investigated, it may come to a verdict that 
those ``faults" will never be activated/excitated during a restricted execution starting 
from legal initial states and with legal inputs. Thus we will call those ``faults" as 
{\bf spurious faults}, since they will not affect the circuit's normal behavior.

Almost all practical sequential logic components can be modeled as finite state machines (FSMs). 
If we apply constrains upon the machine to make it start from designated initial states and 
take in specific legal inputs, a set of reachable states can be derived. 
As long as the ``faults" can be modeled as ``bad states", we can judge whether they are 
spurious faults by checking if they sit in the reachable states. From the spurious fault validation 
perspective, reachability analysis is a must when developing full set of sequential circuit verification
techniques.

There are quite a few methods to perform reachability checking on FSMs. One among them is 
state space traversal. Conventionally the algorithm is based on bit-level techniques such as
binary decision diagrams (BDDs) and Boolean logic. We propose a new traversal algorithm on word-level,
which bring critical advantages. In this chapter the approach will be described and discussed in depth, 
with examples and experiments showing its feasibility when applied on general circuit benchmarks.

\section{Motivation}
The inspiration of this research mainly comes from a journal paper \cite{KallaPartialScan}. 
In that paper, the author proposed an traversal algorithm using concept ``implicit state enumeration".
Concretely, the algorithm is written as follows:

\begin{algorithm}[H]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Sequential circuit, number of registers to scan}
 \KwOut{Scan registers listed in decreasing order of their non-controllability}

  $from^0 = reached = S^0$\;
  $i = 0$\;
  \While{TRUE}
  {
  	$i++$\;
	$to^i = $IMAGE$(\Delta,from^{i-1})$\;
	$new^i = to^i \cdot \overline{reached}$\;
	\For{each state variable $r_j$}
	{
		record\_if\_transitions\_present\_or\_missing$(r_j,new^i)$\;
		compute\_degree\_of\_unsettability$(r_j,new^i)$\;
	}
	\If{$new^i == 0$}
	{
		break\;
	}
	$reached = reached + new^i$\;
	$from^i = new^i$\;
  }\
  \tcc{\ \ \ \ \ \ \ \ \ \ \ \ \ \ FSM traversal completed}\
  \For{each state variable $r_j$}
  {
	\If{missing transition for $r_j$}
	{
		scan state variable $r_j$\;
	}
  }
  $illegal\_states = $bdd\_complement$(reached)$\;
  \For{each state variable $r_j$}
  {
	compute\_degree\_of\_unateness$(r_j,illegal\_states)$\;
	non-controllability$(r_j)=$degree\_of\_unsettability$(r_j)+$degree\_of\_unateness$(r_j)$\;
  }
  order state variables in terms of their non-controllabilities;\tcc{\ Sorting}\
  output the required scan registers\;
%\Return{$from_k(R_{final})$}
\caption {SIMPSON: Scan Register Selection using Implicit State Enumeration\cite{KallaPartialScan}}
\label{alg:SIMPSON}
\end{algorithm}
\DecMargin{1em}