### 2016-10-24 dissertation/abstract.tex
dissertation: milestone 1: first draft

\begin{center}{\bf ABSTRACT}\end{center}

Verification of sequential circuits remains to be a topic of all time. With increasing size of new
integrated circuits, sequential circuit designers may face much more complicated problems on 
logic errors, unexpected delays and failures on critical path. 
This dissertation addresses the problem of sequential circuit verification at the word-level
and is based on concepts derived from algebraic geometry.

Analyzing the sequential circuits at word level is an efficient way
of {\it abstraction}, which may lower the complexity of verification by efficient representation of the
state-space. We manage to model the verification properties and the gate-level sequential circuit 
implementations over Galois fields of the type ${\mathbb{F}}_{2^k}$ by means of polynomial ideals and 
their canonical representations --- Gr\"obner bases --- at the level of $k$-bit words. Subsequently, 
techniques from algebraic geometry can be used to reason about the state-space on high-level.
In algebraic geometry, ``bad" states are modeled as varieties while whole system is described using 
polynomial ideals. Without actually solving the system, the states in state space can be investigated 
by manipulating these ideals.

We propose to apply these techniques to traverse a finite state machine (FSM) for reachability 
analysis at the word-level, and also to implicitly unroll sequential arithmetic circuits to 
verify their function. Moreover, as unsatisfiable (UNSAT) cores play an important role in modern 
abstraction-refinement techniques for verification, we propose to investigate a word-level 
analogue of the UNSAT core problem using the  Gr\"obner basis algorithm. The dissertation will 
not only derive new algorithms and techniques, but will also consider efficient CAD 
implementations to target sequential equivalence and model checking problems. 
% \end{abstract}