### 2016-10-24 dissertation/conclude.tex

\chapter{Conclusions and Future Work}
\label{ch:conclude}

\section{Future Work}
\subsection{Multivariate polynomial ideal based FSM Traversal}
\subsection{New Diagram Structure accelerating Polynomial Reduction}
\subsection{Interpolation extraction using GB Algorithm}
\subsection{Verification of Integer Arithmetic Circuits}