### 2016-10-24 dissertation/prev.tex

\chapter{Previous Work}
\label{ch:prev}
\section{SAT,DD,AIG-based Sequential Circuit Verification}

\section{Bounded Model Checking}

\section{Model Checking with Abstraction Refinement}

\section{Word-level Techniques in Sequential Circuit Verification}
Term rewriting?

\section{Verification of Galois Field Circuits}

\section{Conventional UNSAT Core Extraction}

\section{Conclusion Remarks}