### 2016-11-09 dissertation/intro.tex
dissertation v2_chap 3 4_update1

\chapter{Introduction}
\label{ch:intro}
\section{Design, Test and Verification of Modern Sequential Circuit Components}

\section{Formal Verification Techniques}
\subsection{Conventional Formal Verification}
\subsection{Word-level Verification Techniques}

\section{Objective, Motivation and Contribution of this Dissertation}
\subsection{Word-level Reachability and FSM Traversal}
\subsection{Application to Galois Field Multipliers}
\subsection{Abstractions for Sequential Circuits}
\subsection{Other Applications}

\section{Dissertation Organization}