### 2016-11-09 dissertation/prelim_GF.tex

\chapter{Galois Field and Sequential Arithmetic Circuits}