### 2016-12-01 dissertation/intro.tex
dissertation v2: part intro update 1

\chapter{Introduction}
\label{ch:intro}
With the increasing size of integrated
circuits, sequential circuit designers face complicated problems of
design errors in specification models and implementations. These
errors are usually modeled as ``bad" states, and the
circuits/functional components are modeled as finite state machines
(FSMs). Once state reachability is analyzed, the existence of errors
can be identified by checking whether ``bad" states are {\it
  reachable} from certain initial states. Temporal logic {\it model
  checking} (MC) formulations and solvers are often used for this
purpose. Once the designs and specification models are validated using
model checking, optimized implementations of sequential circuits are
synthesized. A subsequent problem then needs to be solved to prove
that the sequential circuit implementations are equivalent to the
original specification models, i.e. {\it Sequential Equivalence
  Checking   (SEC)}.

Reachability analysis forms the backbone of most sequential
verification techniques. As the state-space of FSMs increases,
reachability analysis forms a fundamental bottleneck in sequential
verification. Contemporary approaches employ various techniques to
overcome this state-explosion problem: i) Bounded model checking
\cite{bitlevel1} traverses the FSMs for a fixed number of steps $k$
($k$-BMC) to check whether a property violation can occur in $k$ or
fewer steps. ii) Analyze over-approximations (or {\it abstractions})
of the state-space. Abstraction proves properties on the system by
first simplifying it, and when the abstraction does not satisfy the
same properties as the original one, a process of refinement is
needed. {\it Craig interpolants} \cite{craig-interpolate}
\cite{pudlak:ci} are used as a means of abstraction and reachability
analysis for linear temporal logic (LTL) model checking
\cite{mcmillan:interpolation}. Similarly, counterexample guided
abstraction refinement (CEGAR) \cite{cegar-journal} uses proofs of
unsatisfiability (UNSAT cores) to refine the abstractions. iii) the
recent breakthrough method of \cite{bradley:pdr} where the set of
over-approximations to forward reachable states are refined with
inductive constraints -- property directed reachability (PDR).

While the above techniques have made significant strides in sequential
verification, numerous practical instances remain unsolved. One issue
with all of the above techniques is that they mostly use bit-level
constraints to model the transition relations and sets of 
states. We have access to instances from our industrial
collaborators (Calypto Design Systems, now Mentor Graphics) where
the designs are expressed at the level of bit-vector words
(e.g. Matlab code, Verilog), and these word-level abstractions are
rarely exploited in verification. The problem is further exacerbated
when there are arithmetic operators on word-level operands embedded in
the control logic. While attempts have been made towards word-level
predicate abstraction \cite{Jain05wordlevel} \cite{mcmillan:cav06}
\cite{mcmillan:impact}, {\it using a purely word-level representation
  of the state-space, the properties and their abstractions have not
  been fully explored as another dimension in improving sequential
  verification.}

There are several advantages in exploiting word-level information for
verification. A number of designs have  their
datapaths and/or system-level models described as word-level RTL
models.  Exploiting word-level instead of bit-level information is one
way of abstraction -- a key technique to reduce the state space of a
FSM. It has the effect of combining sets of states with similar 
properties. During reachability analysis, if we use bit-level
variables to  represent the states, the representations may become too
large to handle. However, when a ``bundle" of bit-level variables are
represented as {\it only one word-level variable}, the set of
reachable states can be represented by a  word-level constraint
expression; which may lower verification complexity.

{\bf Overall Research Objective:} In this proposal we investigate a
set of new, promising approaches for {\it word-level representation,
reachability analysis and abstraction} with applications to SEC,
$k$-BMC and abstraction-based unbounded LTL model checking
(A-UMC). Our techniques operate at the word-level and they are based
largely on the concepts from {\it algebraic geometry}.

For word-level SEC, we are given two designs, or their corresponding
Mealy/Moore FSMs ${\mathcal{M}}_1,{\mathcal{M}}_2$, along with their
initial starting states $S_0^1,S_0^2$. We wish to prove the absence of
a sequence of inputs (string) that distinguishes the initial
states \cite{coudert:iccad90}
\cite{coudert1990verification}. Fundamentally, this requires 
the construction of a product machine; and the main research problem
relates to that of performing {\it FSM traversal \cite{touati:iccad90}
but  at the word-level}. Analogously, in the case of MC, the problem
is setup w.r.t. a FSM $\mathcal{M}$, a set of initial states $S_0$ and
a set of property states $p$. Techniques are to be researched that
verify that there exist no sequence of transitions from an initial
state to a non-property state (``bad'' state). These problems have to
be solved in the context of word-level verification -- i.e. data
representation, abstraction (Craig interpolation, UNSAT cores) and
algorithm execution has to be carried out at word-level.

{\bf The Algebraic Geometry Model and Rationale:} We will represent
the FSMs -- the transition relations -- by means of a set of
multi-variate polynomials with coefficients from the finite (Galois)
field $\Fkk$ of $2^k$ elements, i.e. polynomials in the ring
$\Fkk[x_1,\dots,x_d]$. Each state of a FSM is identified with a
Boolean assignment to a set of $k$-bit state register variables
$S=\{s_0,\dots,s_{k-1}\}$. Therefore, we can consider each ($k$-bit)
state as a word-level element $S$ of the finite field
$\Fkk$. Algorithms can directly operate on polynomials in word-level
variable $S$.

Boolean functions with $k$-bit inputs and $k$-bit outputs 
$f: \B^k \rightarrow \B^k,\B = \{0, 1\}$ can be construed as functions
$f: \Fkk \rightarrow \Fkk$. It is well-known that over the finite
field ($\Fq$) of  $q$ elements, {\it every function} $f: \Fq
\rightarrow \Fq$ is a polynomial function \cite{ff:1997}. Moreover,
there exists a unique canonical polynomial $\FF$ that describes $f$.
This implies that one can derive a {\it  canonical, polynomial
  abstraction} of the function as $Z = \FF(A)$ where $Z, A$ are
word-level symbols representing $k$-bit operands. The concept also
generalizes to functions with different input/output bit-vector sizes,
i.e. functions of the type $f: \B^n \rightarrow \B^m$, modeled as
polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow
{\mathbb{F}}_{2^k}$, where $k=LCM(n,m)$ \cite{ff:1997}.  
This implies that the FSM's transition relations can be
represented as polynomial functions (ideals) in $\Fkk$, and values of
state variables can be represented as solutions to these polynomials
(variety of the ideal). Subsequently, the ideal-variety correspondences
in algebraic geometry can be applied to implement symbolic reasoning
about state reachability. Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
our model provides a single, unified and bit-precise representation
for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
constraints.

The decision and abstraction procedures in our setting will rely on
the theory and technology of {\it \Grobner bases (GB)}. GB-based
algebraic reasoning is very powerful; in fact it is known to be
strictly stronger than resolution \cite{CEI:stoc-96}. Therefore, in
light of the above discussion, using concepts from algebraic geometry
and \Grobner bases over $\Fkk$, we can introduce another dimension of
word-level abstraction to the techniques in sequential verification.

{\bf The core focus of this research:} The use of algebraic geometry
for sequential circuit verification and symbolic model checking has
been presented before. Avrunin presented the
concept of symbolic MC using algebraic geometry in
\cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
GB-algorithms for CTL, LTL and 
bounded MC over Boolean rings. However, these approaches are a
straight-forward transformation of the problem to {\it bit-level}
Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
the concepts of word-level reachability, abstraction-refinement using
interpolation or UNSAT cores, etc., that we desire was not the focus of
\cite{Avrunin:CAV} \cite{vardi-iasted07}.

In this research, we will focus on three specific research problems: 
\begin{enumerate}
\item How to perform reachability analysis at the word-level? This
  of course forms the foundation for word-level SEC and MC.
\item What is the analog of Craig interpolation in an algebraic
  geometry setting (theoretical challenge), and how can we compute
  various interpolants using GB-engines efficiently (practical
  implementation)? This will enable word-level abstraction of the
  state-space. 
\item Explore the concept, and the computation, of unsatisfiable (UNSAT)
  cores of  a set of polynomials. Many abstraction-refinement
  techniques make use of UNSAT cores of CNF-formulas in MC. Therefore,
  our investigation of UNSAT cores of polynomial ideals will enable
  similar applications at word-level. 
\end{enumerate}

\section{Design, Test and Verification of Modern Sequential Circuit Components}\section{Hardware Design and Verification Overview}
The typical design flow of a hardware system, 
as shown in Fig. \ref{fig:cadflow}, starts with a hardware system 
specification, which describes the necessary functions and parameters that 
the system must perform and adhere to. The specification is typically
modelled using a transaction-level model (\emph{TLM}), which describes 
communication details between large circuit modules. 
The TLM is then translated 
into a register-transfer-level (\emph{RTL}) description, which is composed of abstracted, 
interconnected circuit blocks that compose the entire system. RTL is 
typically implemented 
in hardware description languages (\emph{HDL}) such as Verilog and VHDL, 
which are the most popular choices in the industry.
Next, the RTL is optimized and converted into a \emph{netlist}, i.e. a large 
collection of small physical blocks (MOSFET, Boolean logic gates, etc.) and 
the inter-connections (wires) between them. 
Lastly, the netlist is further optimized and then mapped onto a physical space 
on a chip, which is then sent off for fabrication. This entire design flow
is automated by Computer-Aided Design (\emph{CAD}) tools.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/designFlow}
}
\caption{Typical hardware design flow.}
\label{fig:cadflow}
\end{figure}
}

%\subsection{The Hardware Verification Imperative}
When moving from one abstraction level of the hardware design process to the 
next, an important issue arises: how can one ensure that the functionality of 
the optimized design matches original spec? 
Bugs in hardware design which are 
not caught early can have costly effects later, such as the need for a 
redesign. Bugs in arithmetic circuits can be especially catastrophic.
One infamous example is the 1994 floating point division (FDIV)
bug that affected the Intel Pentium chip \cite{nicely:FDIV}, 
and subsequently cost the company \$475 million because it was  
discovered after the chip's release. In another more fatal case, during the
Gulf war, an American Patriot Missile battery failed to intercept an incoming
enemy missile due to an arithmetic error \cite{arnold:patriot}.
Since hardware bugs can have significant consequences, 
there has been extensive work in field of hardware verification to find and
eliminate bugs prior to fabrication.

The two main methodologies used in hardware verification are simulation and 
formal verification. \emph{Simulation} checks correctness by applying exhaustive 
assignments to the circuit inputs and verifying correctness of the output. 
This ensures that the circuit performs as designed under all possible 
inputs. Such exhaustive testing is quite effective for smaller circuits. 
However, as the size of the circuit increases, it becomes 
computationally infeasible to simulate all possible test vectors. This is the 
case with Galois field arithmetic circuits, which are commonly very large in 
real-world applications. Often for such large circuits, simulations of a smaller and more 
manageable subset of test vectors are employed to catch bugs. While these tests
can increase confidence in the correctness of the design,  
{\it they do not guarantee correctness} since every data-flow of the design hasn't
been analyzed.
%Such is the case with large Galois field arithmetic circuits, so we instead 
%focus on formal verification.

\section{Formal Verification}
Instead of simulating input vectors, \emph{formal verification} utilizes 
mathematical theory to reason about the correctness of hardware designs.
%which overcomes some limitations of simulation. 
Formal verification has two main forms: property checking and equivalence 
checking.

{\it Property checking} (or property verification) verifies
that a design satisfies certain given properties. Property checking is done mainly 
in the form of theorem proving, model checking, or approaches which 
combine the two.
\begin{enumerate}
\item \emph{Theorem proving} \cite{theoremproving:91} requires the existence of
mathematical descriptors of the specification and implementation of the 
circuit. Theorem provers apply mathematical rules to these descriptors to
derive new properties of the specification. In this way, the tool can reduce
a proof goal to simpler sub-goals, which can be automatically verified.
However, generating the initial proof-goal requires extensive guidance from
the user, so there is an overall lack of automation in theorem 
proving.
\item \emph{Model checking} \cite{modelcheck:99} is an approach
to verifying finite-state systems where specification 
properties are modeled as a system of 
logic formulas. The design is then traversed to check if the 
properties hold. If the design is found to violate a
particular property, a counter-example is generated which exercises the
incorrect behavior in the design. Such counter-examples allow the designer
to trace the behavior and find where the error in the design lies.
Modern model checking techniques use the result to automatically refine
the system and perform further checking.
These tools are typically automated, and thus have found widespread 
use in CAD tool suites.
\end{enumerate}

{\it Equivalence Checking} verifies that two different representations of
a circuit design have equivalent functionality. An example of equivalence
checking as it applies to the hardware design flow is shown in
Fig. \ref{fig:equivflow}.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/designVerification}
}
\caption{Equivalence checking as applied to the hardware design flow.}
\label{fig:equivflow}
\end{figure}
}

There are two major
equivalence checking techniques: graph-based
and satisfiability-based.
\begin{enumerate}
\item \emph{Graph-based} techniques construct a canonical graph 
representation, such as a Binary Decision Diagram (\emph{BDD}) or one of
its many variants, of each circuit. A linear comparison is then conducted to 
determine whether the two graphs are isomorphic. Since the graph 
representation is canonical, the graphs of the two circuits will be 
equivalent if and only if the circuits perform the same function.
\item \emph{Satisfiability} techniques construct a miter of the two circuits,
typically in a graph such as an And-Inverter graph (\emph{AIG}). A
\emph{miter} is a combination of the two circuits with one bit-level output, which 
is only in a "1" state when the outputs of the circuits differ given 
the same given 
input, as shown in Fig. \ref{fig:miter}. 
A satisfiability (\emph{SAT}) tool \cite{csat} 
is then employed to simplify the graph and find a solution to the miter, 
i.e. find an input for which the 
miter output is "1". If a solution is found, this solution acts as a 
counter-example of when the circuit outputs differ; otherwise the circuits
are functionally equivalent.
\end{enumerate}

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.8\textwidth]{./figures/betterMiter}
}
\caption{A miter of two circuits.}
\label{fig:miter}
\end{figure}
}

%\subsection{Computer-Algebra Based Formal Verification}
Certain formal verification methods use \emph{computer-algebra} and \emph{algebraic
geometry} techniques based on mathematical theories.
Unlike SAT-based verification, modern algebraic geometry 
techniques do not explicitly solve the constraints to find a solution; 
rather, they reason about the presence or absence of solutions, or explore
the geometry of the solutions.
These methods \cite{Avrunin:CAV} \cite{condrat-tacas07} \cite{gbverify:2007} 
transform the circuit design into a polynomial system. Typically, this system
of polynomials is then used to compute a Gr\"obner basis \cite{gb_book}. 
Computation of Gr\"obner bases allows for 
the easy deduction of important properties of a polynomial system, 
such as the presence or absence of 
solutions. These properties are then leveraged to perform 
verification. Unfortunately, such a computation 
has been shown to be doubly 
exponential in the worst case, and thus these methods have not been 
practical for real-world applications. However, recent
breakthroughs in computer-algebra hardware verification have shown
that it is possible to overcome the complexity of this computation while
still utilizing the beneficial properties of a Gr\"obner bases
\cite{lv:phd}.

\section{Importance of Word-level Abstraction}
Most formal verification techniques can benefit from word-level abstractions 
of the circuits they verify.
Abstraction is defined as state-space reduction, i.e{\text . }abstraction
reduces state-space by mapping the set of states of a system to a smaller 
set of states. Because the new representation contains fewer states, it
is easier to comprehend and thus easier to use. 
Word-level abstraction focuses specifically on abstracting a word-level
representation of a circuit out of a bit-level representation. For example,
a bit-level representation of an integer multiplier is represented by a
collection of Boolean inputs and outputs, whereas a word-level
abstraction hides the underlying logic and represents the circuit as two 
integer inputs and one integer output, e.g. $Z=A\cdot B$. As the bit-size of the
multiplier increases, the logical implementation of the multiplier grows (typically
exponentially) while the word-level abstraction stays the same.

Word-level abstractions have a wide variety applications in formal 
verification. Theorem proving techniques can leverage abstraction as an 
automatic decision procedure or as a canonical reduction 
engine. For example, since RTL is composed of circuit blocks that represent 
the underlying circuit, RTL verification methods can exploit 
abstractions of these blocks.
This is seen in the following RTL verification methods:
\begin{itemize}  
\item Model checking \cite{kroening:model}, 
where an approximation abstraction of RTL blocks is generated and then 
refined.
\item Graph-based equivalence 
checking \cite{WLS} \cite{arditi:bmd}, where abstraction methods are used
to generate a canonical word-level graph representation of the circuit.
\item Satisfiability-based equivalence checking \cite{lpsat}, where 
abstractions are used identify symmetrics and similarities in order to 
minimize the amount of logic that is sent to the 
SAT tool. 
\end{itemize}

Other equivalence checking techniques that employ abstractions 
include satisfiability modulo theory ({\it SMT}) techniques \cite{boolector} \cite{bryant:tacas07}, 
which are similar to SAT except they operate on higher-level data
structures (integers, reals, bit vectors, etc.), as well as 
constraint solving techniques \cite {ms:research} \cite{tew:iccad08}.
In general, RTL equivalence checking approaches would ideally maintain a 
high-level of abstraction while still retaining sufficient lower-level 
functional details  (such as bit-vector size, precision, etc) 
\cite{gupta_survey}.

Word-level hardware abstractions also have applications in RTL and datapath 
synthesis \cite{demicheli:iccad_98} \cite{demicheli:dac_99}
\cite{demicheli:tcad_03}. 
Abstractions of circuits allow for design reuse, which allows for tool-automated 
synthesis of larger circuit blocks.
Since hardware design specifications tend to be word-level, synthesis tools 
can use these larger circuit blocks to generate and optimize the
datapaths and create the RTL of the system. Thus, in order for a circuit to 
be used by these automated synthesis tools, its word-level abstraction must
be known.

Finally, abstractions can also be applied to detect malicious 
modifications to a circuit, potentially inserted as a hardware trojan horse.
Hardware trojans, a relatively new security concern in the hardware 
industry, use certain techniques to add incorrect behavior to a 
design. 
This behavior is only activated under certain rare circumstances that only 
the mal-intent designer has knowledge of.
The behavior is purposely hidden and is very difficult to encounter during 
simulation of the design. A manufactured chip with a subsystem 
that contains a hardware trojan could compromise the entire system in which 
it is used.
In some hardware trojan cases, formal verification techniques may be applied 
to catch a bug in a design and provide a counter-example which exercises it. 
However, it can be difficult to tell whether the bug in the design was 
introduced intentionally of not. On the other hand, word-level abstractions 
of bit-level circuits {\it effectively reverse-engineer the true function 
implemented by the circuit}, which could be used to determine the designer's 
true intention.

\section{Objective, Motivation and Contribution of this Dissertation}
\subsection{Word-level Reachability and FSM Traversal}
\subsection{Application to Galois Field Multipliers}
\subsection{Abstractions for Sequential Circuits}
\subsection{Other Applications}

\section{Dissertation Organization}
{\bf Proposal Organization:} The rest of the proposal is organized as
follows: The following Subsection describes the intellectual merit of
the proposed work.  Section \ref{sec:prelim} covers preliminary
concepts and notation. Section \ref{sec:reach} describes the basic
concept of word-level FSM traversal and the research problems to
investigate. Section \ref{sec:craig} explores the algebraic geometry
analog of Craig Interpolants. Section \ref{sec:core} describes
algorithmic techniques to derive UNSAT cores of polynomial
ideals. Section \ref{sec:plan} puts the whole research in perpective,
by demonstrating  with the help of an example how abstraction via
unsat cores in algebraic geometry can simplify BMC. 
Section \ref{sec:impact} addresses the broader impact of the
proposed activity and concludes the proposal.