### 2016-12-02 dissertation/intro_xiaojun.tex

Following the Moore's law, the level of integration on modern VLSI becomes very high, 
such that a single digital VLSI design can implement a true system-level component.
The design process also evolves from manual design with little validation, to 
a formal 3-step procudure which needs collaboration of team with large number of 
engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
2) Implement, which is to refine the design through all phases with the assistance of Computer-Aided-Design (CAD)
tools; 3) Verify, which is to verify the correctness of design and implementation.

Nowadays the verify step is usually completed by a team with specialities on test, verification and validation of 
circuits. This step is also automated as an indispensable part of CAD tools, when circuit synthesis is performed. 
Figure \ref{fig:designflow} shows the typical synthesis flow, which covers all procudures starting from 
register-transfer-level (RTL) description (using hardware design languages, i.e. HDL) to  the 
physical design on silicon (depicted by the layout). The objective of verification in synthesis is 
to ensure the implementation is consistent with the original design intent. Verification is 
an important quality control measure before sending design layout to the VLSI foundaries.
Considering the high cost of fabrication, any faults and errors in the design will bring considerable 
waste of funds for the designers. On the other hand, all aspects of the society increasingly depend on 
the stability and accuracy of digital VLSI circuits, even small flaws or short-time failures can cause 
huge loss, especially in medical applications, military facilities and financial systems.
Therefore, it is of utmost importance to verify the correctness of VLSI designs.

One way to perform the verification is called {\it simulation}. It is defined as the collage of all circuit validation 
methods which apply stimulations on the inputs of circuit model and verify correctness of the outputs.
However, simulation is not a complete solution to circuit verification problems. In modern designs with 
large number of logic components and complicated architecures, it is impractical to simulate all possible 
test vectors. Usually only test vectors that correspond to typical failures are selected in simulation, which 
cannot cover unexpected failure patterns caused by special inputs. The notorious Intel FDIV bug \cite{nicely:FDIV}
is a good example when simulation fails. Failure occurs when only 5 entries in an $84\times 16$ look-up table
(LUT) are activated,  which is rarely used in most divisions. Because of the limitation of simulation,  
test engineers from Intel failed to detect the bug,  which brought $\$475$ million dollors recalling bill
for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.

Formal verification can provide $100\%$ fault coverage from 2 aspects. On the one hand,  it formalize properties
for the circuit model which are irrelevant to specific input signals,  and prove them mathematically.
On the other hand,  it adopts formal languages to strictly describe the design intents and detailed implementation, 
and deduces circuit function from the implementation. These descriptions are named as {\it specifications}.