### 2016-12-13 dissertation/slides/defense.tex
dissertation: slides v1

% Lump tex of Xiaojun's defense slides

\documentclass[xcolor=dvipsnames]{beamer}
%%\usetheme{default}
\usecolortheme[named=Maroon]{structure}
%\usetheme{Boadilla}
\usetheme{Madrid}
\useoutertheme{default}%[footline=empty]{infolines}

% \usepackage{helvet}
% \usepackage{enumerate}
% \usepackage{amsmath}
% \usepackage{amsfonts}
% \usepackage{graphicx}
% \usepackage{ulem}
\usepackage{multirow}
\usepackage{comment}
\usepackage{xspace}
%\usepackage{columns}

\usepackage[absolute,overlay]{textpos}
\usepackage[ruled,linesnumbered]{../algorithm2e}
%%for algorithm2e package, label has to be following caption in the same line!!!
\renewcommand{\algorithmcfname}{ALGORITHM}
\renewcommand{\thealgocf}{}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\SetAlCapHSkip{0pt}
\IncMargin{-\parindent}

%\RequirePackage{algorithmic}
%\RequirePackage{algorithm}
% \renewcommand{\algorithmicrequire}{\textbf{Inputs:}}
% \renewcommand{\algorithmicensure}{\textbf{Outputs:}}

%\newtheorem{theorem}{Theorem}
%\newtheorem{lemma}{lemma}
%\newtheorem{corollary}{Corollary}
%\newtheorem{proposition}{Proposition}
%\newtheorem{Q}{Question}
%\newtheorem{Exa}{Example}
%\newtheorem{Definition}{Definition}

\newcommand{\Fq}{{\mathbb{F}}_{q}}
\newcommand{\Fp}{{\mathbb{F}}_{p}}
\newcommand{\Fpk}{{\mathbb{F}}_{p^k}}
\newcommand{\Fkk}{{\mathbb{F}}_{2^k}}
\newcommand{\Z}{{\mathbb{Z}}}
\newcommand{\Zkk}{{\mathbb{Z}}_{2^k}}
\newcommand{\Fkkx}[1][x]{\ensuremath{\mathbb{F}}_{2^k}[#1]\xspace}
\newcommand{\Grobner}{Gr\"{o}bner\xspace}
%\newcommand{\Grobner}{Gr\"{o}bner}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\Func}{{\mathcal{F}}}
\newcommand{\N}{{\mathcal{N}}}
\newcommand{\G}{{\mathcal{G}}}
\newcommand{\F}{{\mathbb{F}}}
\newcommand{\B}{{\mathbb{B}}}
\newcommand{\C}{{\mathbb{C}}}
\newcommand{\R}{{\mathbb{R}}}

\title[PhD Dissertation Defense]{Word-Level Abstractions for Sequential Design Verification using Algebraic Geometry}

\author[Xiaojun Sun]{{\bf Xiaojun Sun}, PhD Candidate}

%\email{rostamian@umbc.edu}
\institute[Univ. of Utah]{
\includegraphics[height=22mm]{./Ulogo_1200px.png}\\
\    \\
\    \\
\    \\
Electrical and Computer Engineering, University of Utah\\
\    \\
Advisor: {\it Priyank Kalla}
}

\date{Dec 16, 2016}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%----------- titlepage ----------------------------------------------%
\begin{frame}[plain]
  \titlepage

\end{frame}

%\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Outline}}
\bi
\item Contributions
\item Motivations
\item Previous Work
\item Preliminaries
	\bi
	\item Finite fields
	\item Polynomial algebra \& Algebraic geometry
	\item Projection
	\ei
\item Projection based abstraction 
	\bi 
	\item Application: \alert{Reachability analysis}
	\item Application: \alert{Sequential arithmetic ckt verification}
	\ei
\item UNSAT core based abstraction
	\bi
	\item \alert{UNSAT core extraction using Gr\"obner basis refutation}
	\item Application: Bounded model checking (BMC) with abstraction refinement
	\ei
\item Conclusion \& Future work
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Contributions}}
\bi
\item Word-level reachability analysis -- analog if implicit state enumeration
	\bi 
	\item {\bf Q:} Why word-level? 
	\pause
	\item {\bf A:} Datapath $\gets$ word-level info 
	\pause
	\item {\bf A:} Simplify representation (abstraction) of state-space $\to$ Efficiency!
	\ei
\pause
\item Apply word-level reachability algo. to sequential arithmetic circuit verification
	\bi
	\item Abstraction $\to$ word-level signature each time-frame
	\item Word-level unrolling
	\ei
\pause
\item UNSAT cores in Algebraic geometry
	\bi
	\item Analyze Buchberger's algorithm $\to$ extract refutation proof $\to$ UNSAT core
	\item Structure of refutation proof $\to$ refine UNSAT core
	\item Apply UNSAT core to $k$-BMC abstraction refinement
	\ei
\pause
\item Implement above algos: C++ \& \textsc{Singular}
	\bi
	\item For sequential GF multipliers: overwhelmingly better than contemporary tools
	\ei
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = linkSTG]{\large{Motivation I: BFS state space traversal}}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{./stg_fig.eps}
\caption{State Transition Graph}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\bi
\item Initial state: $\{00\}$
\item Iteration 1: 
	\bi
	\item Start from $\{00\}$
	\item One-step transition: $\{01,10\}$
	\item Newly reached: $\{01,10\}$
	\ei
\item Iteration 2: 
	\bi
	\item Start from $\{01,10\}$
	\item One-step transition: $\{00,10\}$
	\item Newly reached: $\emptyset$
	\ei
\item All reachable states detected. Final reached states: $\{00,01,10\}$
\pause
\item Still need bit-level Boolean variables to represent states
\ei
\end{column}
\end{columns}
\hyperlink{pptpage2}{\beamergotobutton{Go back to example}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = motiv2]{\large{Motivation II: Sequential arithmetic circuits verification}}
\vspace{-0.1in}
\begin{figure}[H]
\centering{
\includegraphics[width=4.5in]{../newfig/convention.eps}
}
\end{figure}
\vspace{-0.2in}
\bi
\item Problem: Verify the function of sequential arithmetic circuit
	\bi
	\item Operands preloaded into registers
	\item After $k$ clock cycles, give desired output
	\ei
\item Conventional: explicitly unroll $k$ time-frames (bit-level) and setup miter
	\bi
	\item Check miter output: SAT, BDDs, AIGs
	\item Bit-blasting!
	\ei
\ei
\hyperlink{Moore}{\beamergotobutton{Go back to Moore machine}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = motiv3]{\large{Motivation III: $k$-BMC with abstraction refinement}}
\IncMargin{1em}
\begin{algorithm}[H]
\SetAlgoNoLine
	\KwIn{$M$ is the original machine, $p$ is the property to check, $k$ is the number of steps unrolling $M$}
  $k = $ InitValue\;
  \eIf{$k$-BMC$(M,p,k)$ is \textbf{SAT} (reachable states after $k$ steps violates $p$)}
  {
	\Return{``Found error trace"}
  }
  {
	\alert{Extract UNSAT core $\mathcal P$ of $k$-BMC} \;
	$M' = $ \textit{ABSTRACT}$(M,\mathcal P)$\;
  }
  \eIf{MODEL-CHECK$(M',p)$ returns \textbf{PASS}}
  {
	\Return{``Passing property"}
  }
  {
	Increase bound $k$\;
	goto Line 2\;
  }
\caption{$k$-BMC with Abstraction Refinement (L. Zhang'05)}
\end{algorithm}
\DecMargin{1em}
\hyperlink{refine}{\beamergotobutton{Go back to Abs refine}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Previous work}}
\bi
\item Sequential Equivalence Checking (SEC): bit-blasting, or structural info dependency
	\bi
	\item Sequential miter
	\item Unroll, then use DDs, SAT or AIGs (Combinational)
	\ei
\item Symbolic model checking (counterexample, IC3): SAT/BDDs in nature
\item Word-level techniques (term rewriting, uninterpreted function): ?
\item Algebraic geometry methods
	\bi
	\item Gr\"ober basis in model checking [Avrunin,CAV'96;Vardi,IASTED'07]: Analog of bit-level Boolean functions
	\item Abstract word-level polynomial representation for arbitrary combinational ckt [Pruss,TCAD'16]
	\ei
\pause
\item No purely word-level sequential verification: data/abstraction/algorithm
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Finite/Galois Fields}}
\textbf {Galois field} $\mathbb{F}_q$ is a finite field with $q$
elements, $q = p^k$, $p = $prime
\begin{itemize}
\item 0,1 elements, commutative, associate, distributive laws
\item Closure property: $+,-,\times$, inverse ($\div$)
\end{itemize}

Our interest: $\mathbb{F}_{q} = \Fkk$ ($q = 2^k$)
\begin{itemize}
\item  $\mathbb{F}_{2^k}$: $k$-dimensional extension of  $\mathbb{F}_{2}=\{0,1\}$
	\begin{itemize}
	\item $k$-bit bit-vector, AND/XOR arithmetic
	\end{itemize}
\end{itemize}

To construct $\mathbb{F}_{2^k}$
\begin{itemize}
\item $\mathbb{F}_{2^k} \equiv \mathbb{F}_{2}[x] \pmod {P(x)}$
\item $P(x) \in \mathbb{F}_{2}[x]$, irreducible polynomial of degree
  $k$
\item Operations performed (mod $P(x)$) and coefficient reduced (mod 2)
\bi
\item E.g. $\C   = \R[x] \pmod{ x^2+1}$
\ei
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{Preliminaries: Field construction of $\mathbb{F}_8$}
%\ptsize{10}

Consider: $\mathbb{F}_{2^3}= \mathbb{F}_{2}[x] \pmod {x^3 + x +
  1}$\\ $A \in \mathbb{F}_{2}[x]$ \\ 
$A \pmod {x^3 + x + 1} = a_2 x^2 + a_1 x + a_0$. Let $P(\alpha) = 0$:

\begin{itemize}
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 0, 0\rangle = 0$
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 0, 1\rangle = 1$
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 1, 0\rangle = \alpha$
\item $\langle a_2, a_1, a_0 \rangle = \langle 0, 1, 1\rangle = \alpha
  + 1$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 0, 0\rangle = \alpha^2$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 0, 1\rangle =
  \alpha^2 + 1$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 1, 0\rangle =
  \alpha^2 + \alpha$
\item $\langle a_2, a_1, a_0 \rangle = \langle 1, 1, 1\rangle =
  \alpha^2 + \alpha + 1$
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Polynomial function $f:\Fq\to\Fq$}}

\begin{Theorem}[Fermat's Little Theorem over $\Fq$]
Let $\alpha \in \Fq$, then $\alpha^q = \alpha$. Therefore, $x^q -x$
vanishes on all points in $\Fq$. 
\end{Theorem}
\begin{table}[H]
\label{tab:truthtable}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$}
\end{table}
\pause
\bi
\item A polynomial function representation always exists in $\Fq[A]$ (Lagrange's interpolation)
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Computer algebra terminology}}
\vspace{-0.2in}
%\ptsize{10}
Let $\mathbb{F}_q = GF(2^k)$, and $\overline{\Fq}$ be its closure
\begin{itemize}
\item $\mathbb{F}_q[x_1, \ldots, x_n]$: ring of all polynomials with
  coefficients in $\mathbb{F}_q$

\item Polynomial $f = c_1X_1 + c_2 X_2 + \dots + c_t X_t$
\bi
\item A monomial ordering is imposed on $f: X_1 > X_2 > \dots > X_t$
\item \alert{Leading term} $lt(f) = c_1 X_1, ~~tail(f) = c_2 X_2 + \dots + c_t X_t$
\item Leading coefficient $lt(f) = c_1$ and leading monomial $lm(f) = X_1$
\ei
\begin{itemize}
\item LEX $x> y> z$: ~~$f = \alert{ -2x^3} + 2x^2yz + 3xy^3$
\item  DEGLEX $x>y>z$:  ~~$f = \alert{ 2x^2yz} + 3xy^3 -2x^3$
\item DEGREVLEX $x>y>z$: ~~$f = \alert{3xy^3} + 2x^2yz - 2x^3$
\end{itemize}
\end{itemize}
\bi
\item Leading terms $lt(f)$ play an important role
\bi
\item Affect division results!
\ei
\ei

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Polynomial division}}
Copy page 12 from FMCAD
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Algebraic geometry terminology (cont.)}}
Let $\mathbb{F}_q = GF(2^k)$:
\begin{itemize}
\item $\mathbb{F}_q[x_1, \ldots, x_n]$: ring of all polynomials with
  coefficients in $\mathbb{F}_q$ 
\item Given a set of polynomials:
\begin{itemize}
\item $f, f_1, f_2, \dots , f_s \in \mathbb{F}_q[x_1, \dots, x_n]$
\item Find solutions to $f_1 = f_2 = \dots = f_s = 0$
\end{itemize}
\item \alert{Variety:} Set of ALL solutions to a given system of polynomial equations: $V(f_1, \dots, f_s)$
	\begin{itemize}
	\item In $\mathbb{R}\left[x,y\right]$, $V(x^2+y^2-1)=\{all\  points\  on\ circle: x^2+y^2-1=0\}$
	\item In $\mathbb{R}[x]$, $V(x^2+1)=\emptyset$
	\item In $\mathbb{C}[x]$, $V(x^2+1)=\{(\pm i)\}$
	\end{itemize}
\item Variety depends on the \alert{ideal} generated by the polynomials.
\item Reason about the Variety by analyzing the Ideals
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Ideals \& Gr\"obner bases}}
\vspace{-0.1in}

\begin{Definition}
{\bf Ideals of Polynomials:} Let $f_1, f_2, \ldots, f_s \in
\mathbb{F}_q[x_1, \dots, x_n]$. Let 
\begin{eqnarray}
J = \langle f_1, f_2 \ldots, f_s\rangle = \{f_1 h_1 + f_2 h_2 + \dots + f_s h_s\} \nonumber 
\end{eqnarray}

$J = \langle f_1, f_2 \ldots, f_s\rangle$ is an ideal generated by
$f_1, \ldots, f_s$ and the polynomials are called the generators. 
\end{Definition}

\vspace{-0.1in}
%\ptsize{10}

\begin{itemize}
\item Different generators can generate the same ideal
\item $\langle f_1,\cdots,f_s \rangle=\cdots=\langle g_1,\cdots,g_t
  \rangle$
\item Some generators are a ``better'' representation of the ideal
\item A {\bf Gr\"obner basis} $G$ is a ``canonical'' representation of an ideal
	\bi
	\item $I = \langle F \rangle = \langle G \rangle$, and $V(F)=V(G)$
	\ei
\item Map: set of states $\to$ \alert{variety} of polynomial ideal
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Preliminaries: Buchberger's algorithm}}
 INPUT : $F = \{f_1, \dots, f_s\}$\\
% \ENSURE : $G = \{g_1,\dots ,g_t\}$\\ %, a Gr\"{o}bner basis \\
 OUTPUT : $G = \{g_1,\dots ,g_t\}$\\ %, a Gr\"{o}bner basis \\
  $G:= F$; \\
  REPEAT\\
  \hspace{0.1in} $G' := G$\\
  \hspace{0.1in} For each pair $\{f, g\}, f \neq g$ in $G'$ DO\\
\hspace{0.2in}  $Spoly(f, g) \stackrel{G'}{\textstyle\longrightarrow}_+
r$ \\
\hspace{0.2in}  IF $r \neq 0$ THEN $G:= G \cup \{r\}$ \\
%\hspace{0.2in}  $G(x):=G(x) / x$
UNTIL $G = G'$
%\end{algorithmic}
%\end{algorithm}
\pause
\bi
\item $Spoly(f,g)=\frac{L}{lt(f)}\cdot f - \frac{L}{lt(g)}\cdot g$\par
$L = \text{LCM}(lm(f), lm(g))$, ~~$lm(f)$: leading monomial of $f$
\item Animation 1...
\pause
\item Inspiration of mapping: set of states $\to$ \alert{variety} of polynomial ideal
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Gr\'obner basis with elimination term order}}
\begin{itemize}
\item Let ideal $I = \langle f_1, f_2, f_3 \rangle$ where
	\begin{itemize}
	\item $f_1 = x^2 + y + z - 1$
	\item $f_2 = x + y^2 + z - 1$
	\item $f_3 = x + y + z^2 - 1$
	\end{itemize}
\item The \Grobner basis of $I$ with \alert{elimination} (LEX) order $(x > y > z)$ is
	\begin{itemize}
	\item $g_1 = x + y + z^2 - 1$
	\item $g_2 = y^2 - y - z^2 + z$
	\item $g_3 = 2yz^2 + z^4 - z^2$
	\item $g_4 = z^6 - 4z^4 + 4z^3 - z^2$
	\end{itemize}
\item Notice that $g_2$ and $g_3$ only contain variables $y$ and $z$
	\begin{itemize}
	\item Eliminates variable $x~\Leftrightarrow  ~\exists_x$ in Boolean formula!
	\end{itemize}
\item Similarly, $g_4$ only contains the variable $z$ and eliminates $x$ and $y$
\pause
\item GB($I_{x,y,z}$)$\cap\Fq[z]\Leftrightarrow$ \alert{projection} on variable $z$!
\end{itemize}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Elimination by projection}}
\vspace{-0.1in}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{../newfig/projection.pdf}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{../newfig/proj_reacha.pdf}
\end{figure}
\end{column}
\end{columns}
\bi
\item Projection of variety from $\F[x_1,x_2,x_3]$ to $\F[x_2,x_3]$
\item Projection of ckt ideal's variety on next state (NS) variables $T$
\pause
\item GB($J_{ckt}+J_0$)$\cap\F_2[T] \implies $Next state polynomial $f(T)$!
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{BFS traversal algorithm}}
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
\alert{	$to^i \gets$Img$(\Delta, from^{i-1})$}\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}
\pause
\bi
\item Image function: $\text{Img}(\Delta, from) = \exists _s ~\exists _x ~[ T(s, x, t)
  \land from ] = \exists _s ~\exists _x ~\bigwedge_
  {i=1}^{n} (t_i
\overline{\oplus } \Delta_i)\land from$
\item In $\mathbb B^k$, image function $\Leftrightarrow \exists$
\item In $\mathbb F_{2^k}$, need to implement \alert{quantifier elimination}
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%(copy 17~24 HLDVT)%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Recall: Breadth-First Traversal Algorithm}}
\only<1>{
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}

\textbf{Implement this algorithm by finding analogs in algebraic geometry}
}
\only<2>{
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$\tcp*{\alert{polynomial ideal}}}
  $from^0 = reached = S^0$\tcp*{\alert{set of states $\Leftrightarrow$ variety of ideal}}
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}

\textbf{Implement this algorithm by finding analogs in algebraic geometry}
}
\only<3>{
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$\tcp*{\alert{polynomial ideal}}}
  $from^0 = reached = S^0$\tcp*{\alert{set of states $\Leftrightarrow$ variety of ideal}}
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\tcp*{\alert{GB of Elim ideal}}
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}

\textbf{Implement this algorithm by finding analogs in algebraic geometry}
}
\only<4>{
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$\tcp*{\alert{polynomial ideal}}}
  $from^0 = reached = S^0$\tcp*{\alert{set of states $\Leftrightarrow$ variety of ideal}}
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\tcp*{\alert{GB of Elim ideal}}
	$new^i \gets to^i \cap \overline{reached}$\tcp*{\alert{ideal quotient \& sum}}
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}

\textbf{Implement this algorithm by finding analogs in algebraic geometry}
}
\only<5>{
\begin{algorithm}[H]
\SetAlgoNoLine
 \KwIn{Transition functions $\Delta$, initial state $S^0$\tcp*{\alert{polynomial ideal}}}
  $from^0 = reached = S^0$\tcp*{\alert{set of states $\Leftrightarrow$ variety of ideal}}
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\tcp*{\alert{GB of Elim ideal}}
	$new^i \gets to^i \cap \overline{reached}$\tcp*{\alert{ideal quotient \& sum}}
  	$reached \gets reached \cup new^i$\tcp*{\alert{ideal product}}
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {Breadth-first Traversal Algorithm}
\end{algorithm}

\textbf{Implement this algorithm by finding analogs in algebraic geometry}
}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Intersection and union in algebraic geometry}}
\begin{Definition}
\label{def:sum}
({\bf Sum/Product of Ideals}) If $I = \langle f_1, \dots, f_r\rangle$ and $J = \langle g_1, \dots, g_s\rangle$ are 
ideals in $\mathbb F[x_1, \dots, x_n]$, then the {\bf sum} of $I$ and $J$ is defined as
$$I + J = \langle f_1, \dots, f_r, g_1, \dots, g_s\rangle$$ And the {\bf product} of $I$ and $J$ is defined
as
\begin{equation}
  I \cdot J = \langle f_ig_j\ |\ 1 \leq i \leq r, 1 \leq j \leq s\rangle \nonumber
  \end{equation}
\end{Definition}
\begin{Theorem}
\label{thm:unionintersect}
If $I$ and $J$ are ideals in $\mathbb F[x_1, \dots, x_n]$, then ${\bf
  V}(I + J) = {\bf V}(I) \bigcap {\bf V}(J)$ and ${\bf V}(I \cdot J) =
{\bf V}(I) \bigcup {\bf V}(J)$. 
\end{Theorem}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Complement set in algebraic geometry}}
\begin{Definition}
({\bf Quotient of Ideals}) If $I$ and $J$ are ideals in $\mathbb
  F[x_1, \dots, x_n]$, then $I:J$ is the set
  \begin{equation}
  \{f \in \mathbb F[x_1, \dots, x_n]\ |\ f\cdot g \in I, \forall g \in J\}\nonumber
  \end{equation}
and is called the {\bf ideal quotient} of $I$ by $J$.
\end{Definition}
\begin{Theorem}
\label{thm:quotient}
Let $J_0$ be an ideal of vanishing polynomials over $\Fkk[x_1,\dots,x_n]$, then 
$${\bf V}(J_0:J) = {\bf V}(J_0) - {\bf V}(J) = \overline{{\bf V}(J)}$$
\end{Theorem}
\bi
\item Given ideal $J$, compute $J'$ s.t. $V(J') = \overline{V(J)}$
\item $\to J'=J_0:J$
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Our proposed algorithm of BFS traversal based on algebraic geometry}}

\IncMargin{1em}
\begin{algorithm}[H]
\SetAlgoNoLine
\Indm
 \KwIn{The circuit's characteristic polynomial ideal $J_{ckt}$,
   initial state polynomial $\F(S)$, and LEX term order: bit-level
   variables $x, s, t$ $>$ PS word $S$ $>$ NS word $T$}
\Indp

  $from^0 = reached = \F(S)$\;
  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
  {
  	$i \gets i + 1$\;
  	$G \gets$GB($\langle J_{ckt},J_0, from^{i-1}\rangle$)\tcp*{This step contains bit-level}
%  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest}
	$\langle to^i\rangle \gets G\cap \Fkk[T]$\tcp*{Only word-level $S,T$ onwards}
%	\tcc{There will be a univariate polynomial in $G$ denoting the
%          set of next states in word-level variable $T$}
	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
%	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
%  	\tcc{Use product of ideals to attain a union of newly reached states and formerly reached states}
	$from^i \gets new^i(S\setminus T)$\;
%	\tcc{Start a new iteration by replacing variable $T$ in newly reached states with current state variable $S$}
  }
%  \tcc{Loop until a fixpoint reached: newly reached state is empty}
\Return{$\langle reached\rangle$}
\caption {Algebraic Geometry based FSM Traversal}
\end{algorithm}
\DecMargin{1em}

\hyperlink{pptpage2}{\beamergotobutton{Go to example page 2}}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Full Blown traversal of example circuit using algebraic geometry}}
\bi
\item Initial state $from^0 = S(\{00\})$
\item {\bf Iteration 1:}Compose an elimination ideal $J$ 
\vspace{-0.2in}\\
\begin{align*}
&f_1: t_0- (xs_0s_1+xs_0+xs_1+x+s_0+s_1+1)\\
&f_2: t_1 - (xs_0+x+s_0s_1+s_0)\\
&f_3: S - s_0 - s_1\alpha\\
&f_4: T - t_0 - t_1\alpha
\end{align*}
$J_{ckt} = \langle f_1,f_2,f_3,f_4\rangle$
\begin{align*}
&f_5: x^2-x\\
&f_6: s_0^2-s_0, f_7: s_1^2-s_1\\
&f_8: t_0^2-t_0, f_9: t_1^2-t_1\\
&f_{10}: S^4-S, f_{11}:T^4-T
\end{align*}
$\ \ \ \ \ \ \ \ \ \ \  \ J_0 = \langle f_5,f_6,\dots,f_{11}\rangle$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = pptpage2]{\large{Full Blown traversal of example circuit using algebraic geometry}}
\bi
\item Elimination term order: 
$$\{x,s_0,s_1,t_0,t_1\}~(all~bits)~>~S~(PS~word)~ >~ T~(NS~word)$$
\item Compute the reduced GB for $J = J_{ckt}+J_0+\langle from^0\rangle$
\item Next state 
$$to^1 = \langle T^2+(\alpha+1)T+\alpha\rangle$$ 
\item Mapping to set of states
$$V(to^1) = \{1,\alpha\} \Leftrightarrow \{01,10\}$$
\item Complement of formerly reached state:
$$\langle T^4-T\rangle:\langle T\rangle = \langle T^3+1\rangle$$ 
\item Mapping to set of states
$$V(\langle T^3+1\rangle) = \{1,\alpha,1+\alpha\} \Leftrightarrow \{01,10,11\}$$
\ei
\hyperlink{linkSTG}{\beamergotobutton{STG}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = pptpage3]{\large{Full Blown traversal of example circuit using algebraic geometry}}
\bi
\item Newly reached states:
$$\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle T^2+(\alpha+1)T+\alpha \rangle~(\{01,10\})$$ 
\item Update current reached states 
$$reach = \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$$ 
\item Mapping to set of states
$$V(reached) = \{0,1,\alpha\} \Leftrightarrow \{00,01,10\}$$
\item Update the present states for next iteration
$$from^1 = \langle S^2+(\alpha+1)S+\alpha\rangle$$
\ei
\hyperlink{linkSTG}{\beamergotobutton{STG}}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Full Blown traversal of example circuit using algebraic geometry}}
\bi
\item Iteration 2:
	\bi
	\item Next state: $to^2 = \langle T^2+\alpha T\rangle~(\{00,10\})$
	\item The complement of $reached$:
	$$\langle T^4-T\rangle:\langle T^3+(\alpha+1)T^2+\alpha T\rangle
= \langle T + 1+\alpha\rangle ~(\{11\})$$ 
	\item Newly reached state: 
$$\langle T^2+\alpha T, T+1+\alpha \rangle = \langle \alert{1}\rangle$$ 
	\ei
\item Algorithm terminates
\item Return value (final reachable states):
$$reached = \langle T^3+(\alpha+1)T^2+\alpha T\rangle$$
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%(End of copy)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Improve complexity using RATO}}
\bi
\item Directly compute GB($J_{ckt}+J_0$) in $\Fq$ is costly ($q^{O(d)}$)
\item GB sensitive to term order$\to$Transform to simpler set for GB?
\pause
\item RATO: reverse topological traverse on ckt structure
\item Benefit: Only one pair of poly with non-relatively-prime leading terms\par
$gcd(lt(f_w),lt(f_g))\neq 1$
\ei
\begin{figure}[H]
\centering{
\includegraphics[width=4.5in]{../newfig/new_stg.eps}
\label{fig:fsm}}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Example of using RATO}}
\bi
\item RATO: LEX with $(t_0,t_1)>(a,b,c,d)>(x,s_0,s_1)>T>S$
\ei
\begin{equation*}
f_1: a+xs_0s_1+xs_0+xs_1+x+s_0s_1+s_0+s_1+1
\end{equation*}
\vspace{-0.8cm}
\begin{alignat*}{2}
f_2: b+s_0s_1 &~~~ f_3: c+x+xs_0 &&~~~f_4: d+s_0s_1+s_0 \\
f_5: \alert{t_0}+ab+a+1 &~~~ f_6: t_1+cd+c+d &&~~~ f_7: \alert{t_0}+t_1\alpha+T 
\end{alignat*}
\bi
\item $Spoly$ reduction gives $T+\Func(inputs)$
\ei
\begin{align}
&Spoly(f_5,f_7) \xrightarrow{J_{ckt}+J_0}_{+}T + s_0 s_1 x+\alpha s_0 s_1 \nonumber\\
&+(1+\alpha)s_0 x+(1+\alpha) s_0+s_1 x+s_1+(1+\alpha) x+1\nonumber
\end{align}
\bi
\item {\bf Q:} How to get rid of bit-level inputs?
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Bit-to-word conversion}}
\bi
\item Objective: find $s_i = \mathcal G(S)$
\item Build system of poly eqn by squaring:
\begin{align*}
\label{eqn:alphamat}
\begin{bmatrix}
S \\
S^2 \\
S^{2^2} \\
\vdots \\
S^{2^{k-1}}
\end{bmatrix}
&=
\begin{bmatrix}
1 & \alpha & \alpha^{2} & \cdots & \alpha^{{k-1}}\\
1 & \alpha^{2} & \alpha^{4} & \cdots & \alpha^{2(k-1)} \\
1 & \alpha^{4} & \alpha^{8} & \cdots & \alpha^{4(k-1)}\\
\vdots & \vdots & \vdots & \ddots & \vdots \\
1 & \alpha^{2^{k-1}} & \alpha^{2\cdot 2^{k-1}} & \cdots & \alpha^{(k-1)\cdot 2^{k-1}}
\end{bmatrix}
\begin{bmatrix}
s_0\\
s_1\\
s_2\\
\vdots\\
s_{k-1}
\end{bmatrix}
\end{align*}
\item Transition function $f_T: T+\F(S,x)$
\item Elimination on ideal $\langle f_T,f_S \rangle+J_0'$ using $S,x > T$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment results: word-level traversal}}
\begin{table}[H]
\centering
\caption{Results of running benchmarks using our tool. 
\small{Parts I to III denote the time taken by polynomial divisions,
  bit-level to word-level abstraction and iterative reachability
  convergence checking part of our approach, respectively.}}
{\small 
\begin{tabular}{|c||c|c|c|c|c|c|}
\hline
\multirow{3}{*}{\centering Benchmark} 
 & \multirow{3}{0.9cm}{\centering \# States}
 & \multirow{3}{1.3cm}{\centering \# iterations}
 & \multicolumn{3}{c|}{\multirow{2}{2.0cm}{\centering Runtime (sec)}}
 & \multirow{3}{1.8cm}{\centering Runtime of VIS (sec)} \\
   & & &\multicolumn{3}{c|}{}& \\
  \cline{4-6}
     & & & I & II & III & \\
\hline
\hline
b01 & 18  & 5  & $<0.01$ & 0.01 & 0.02 & $<0.01$\\
b02 & 8 & 5 & $<0.01$  & 0.01 & $<0.01$ & $<0.01$ \\
b06 & 13 & 4 & $<0.01$ & 0.07 & 5.0 & $<0.01$ \\
s27 & 6 & 2 & $<0.01$ & 0.01 & 0.02  & $<0.01$  \\
s208 & 16 & 16 & $<0.01$ & 0.32 & 2.4 & $<0.01$ \\
s386 & 13  & 3 & 1.0 & 7.6 & 8.2  & $<0.01$ \\
bbara & 10 & 6 & 0.04 & 0.01 & 0.04  & $<0.01$ \\
beecount & 7  & 3  &$<0.01$ & 0.01 & 0.01 & $<0.01$ \\
dk14 & 7  & 2  & 45 & $<0.01$ & 0.08 & $<0.01$\\
donfile & 24 & 3  & 12316 & 0.02 & 1.7  & $<0.01$\\
\hline
\end{tabular}
}
\label{tab:result}  
\end{table} 
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = Moore]{\large{Apply FSM traversal to arithmetic ckts}}
\vspace{-0.1in}
\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=4.5in]{../newfig/Moore.eps}
% \vspace{-0.2in}
%\caption{A simple Moore FSM and its state transition graph}
%\end{minipage}
\label{fig:Moore}}
\end{figure}
\vspace{-0.4in}
\bi
\item Model: restricted Moore finite state machine
	\bi
	\item Some sequential arithmetic circuits will give results after running for $k$ clock cycles
	\item The initial operands are preloaded in register files
	\ei
\item State transitions on this model:
$$R_{k} = Tr(R_{k-1}) = Tr(Tr(\cdots Tr(R_{init})\cdots)) = Tr^k(R_{init})$$
\item Word-level unrolling
\ei
\hyperlink{motiv2}{\beamergotobutton{Recall unrolling}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Galois field multiplier}}
SPEC: $R = A_{init}\cdot B_{init} \pmod {P(\alpha)}$ after $k$ clock cycles
\hspace{-0.3in}\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=5.5in]{./new_multi.eps}
% \vspace{-0.2in}
\caption{A 3-bit RH-SMPO and its Moore FSM model}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}
\bi
\item Projection on NS $R',A',B'$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{GF multiplier verification algo.}}
\begin{algorithm}[H] % note [hbt] will fail in beamer
\SetAlgoNoLine
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	\alert{$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle)$ with ATO}\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit verification}
\end{algorithm}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%(copy from DATE18~21)%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment on 3-bit RH-SMPO}}
\begin{figure}[hbt]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=1.5in]{./RH3.eps}
% \vspace{-0.2in}
%\caption{Gate-level circuit of 3-bit RH-SMPO}
%\end{minipage}
\label{fig:RHmulti}}
\end{figure}
\vspace{-0.2in}
\bi
\item The elimination ideal (first iteration):
\begin{align*}
J = &d_0+b_2\cdot a_2,
c_1+a_0+a_2,
c_2+b_0+b_2,
d_1+c_1\cdot c_2,\\
&e_0+d_0+d_1,
e_2+d_1,
r_0'+r_2+e_0,
r_1'+r_0,
r_2'+r_1+e_2,\\
&A+a_0\beta+a_1\beta^2+a_2\beta^4,
B+b_0\beta+b_1\beta^2+b_2\beta^4,\\
&R+r_0\beta+r_1\beta^2+r_2\beta^4,
R'+r_0'\beta+r_1'\beta^2+r_2'\beta^4;
\end{align*}
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment on 3-bit RH-SMPO(2)}}
\bi
\vspace{-3.45cm}
\item $J_0 = \langle x_i^2 - x_i, X^q - X\rangle$
\item $from_0 = \{R, A_{init}+a_0\beta+a_1\beta^2+a_2\beta^4,
B_{init}+b_0\beta+b_1\beta^2+b_2\beta^4\}$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Basic algorithm to verify the function of sequential GF multipliers}}
\begin{algorithm}[H] % note [hbt] will fail in beamer
\SetAlgoNoLine
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	$G \gets$GB$( \langle J + J_0+ from_{i-1}(R,A,B) \rangle)$ with ATO\;
	\alert{$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$}\;
	\alert{$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$}\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit verification}
\end{algorithm}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Experiment on 3-bit RH-SMPO(2)}}
\bi
\item $J_0 = \langle x_i^2 - x_i, X^q - X\rangle$
\item $from_0 = \{R, A_{init}+a_0\beta+a_1\beta^2+a_2\beta^4,
B_{init}+b_0\beta+b_1\beta^2+b_2\beta^4\}$ ($\beta = \alpha^3$)
\item $to_1 : 
R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
$
\item $from_1 = \{
R'+(\alpha^2) A_{init}^4 B_{init}^4+(\alpha^2+\alpha) A_{init}^4 B_{init}^2+(\alpha^2+\alpha) A_{init}^4 B_{init}+(\alpha^2+\alpha) A_{init}^2 B_{init}^4+(\alpha^2+\alpha+1) A_{init}^2 B_{init}^2+(\alpha^2) A_{init}^2 B_{init}+(\alpha^2+\alpha) A_{init} B_{init}^4+(\alpha^2) A_{init} B_{init}^2
, A_{init}+a_2\alpha^3+a_0\alpha^6+a_1\alpha^{12},
B_{init}+b_2\alpha^3+b_0\alpha^6+b_1\alpha^{12}\}$
\item $\cdots$
\item After 3 iterations: $to_3 = \{ \alert{R'+A_{init}B_{init},}
~A_{init}+a_0'\alpha^3+a_1'\alpha^6+a_2'\alpha^{12},
~B_{init}+b_0'\alpha^3+b_1'\alpha^6+b_2'\alpha^{12}\}$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Improve using RATO}}
\begin{algorithm}[H] % note [hbt] will fail in beamer
\SetAlgoNoLine
 \KwIn{Circuit polynomial ideal $J$, vanishing ideal $J_0$, initial state ideal $R (=0), \mathcal{G}(A_{init}), \mathcal{H}(B_{init})$}

  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	\alert{$f_o \xrightarrow{J + J_0+ from_{i-1}(R,A,B)}_+ f_r$ under RATO} \;
	\alert{$to_i(R',A',B')\gets f_r(\{R',A',B'\}\setminus\{r_0,\dots,r_{k-1},a_0,\dots,a_{k-1},b_0,\dots,b_{k-1}\}$}\;
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit verification}
\end{algorithm}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment result: sequential GF multiplier verification}}
\begin{table}[H]
\centering
\caption{\small Run-time for verification of bug-free RH-SMPO circuits
  for SAT, ABC and BDD based methods. \emph{TO} = timeout 14 hrs}
\label{tbl:equiv}
\begin{tabular}{|c||c|c|c|c|} 
\hline
& \multicolumn{4}{|c|}{Word size of the operands $k$-bits}  \\
\hline
Solver & 11 & 18 & 23 & 33 \\
\hline
\hline
Lingeling & 593  & \emph{TO}  & \emph{TO}  & \emph{TO}\\
\hline
\hline
ABC & 6.24 & \emph{TO} & \emph{TO} & \emph{TO}\\
\hline
\hline
BDD & 0.1 & 11.7 & 1002.4 & \emph{TO}  \\
\hline
\end{tabular}
\label{table:satbdd}  
\end{table} 
\begin{table}[H]
\centering
\label{tab:Cpp}
{\small 
\begin{tabular}{|c|c||c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c||}{\centering Operand size $k$} & 36 & 60 & 81 & 100 & 131 & 162 \\
\hline
\multirow{2}{1.5cm}{\centering RH-SMPO} & \#Polynomials & 4716 & 12960 & 21870 & 35600 & 56592 & 92826 \\
\cline{2-8}
 & Runtime & 14.3 & 213.3 & 1343 & 4685 & 26314 & 124194 \\
\hline
\multirow{2}{1.5cm}{\centering Agnew's SMPO} & \#Polynomials & 2700 & 7380 & 13356 & 20300 & 34715 & 52974 \\
\cline{2-8}
 & Runtime & 10.2 & 212.0 & 2684 & 4686 & 56568 & 119441 \\

 \hline

\end{tabular}
}
\end{table}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[label = refine]{\large{Abstraction refinement}}
\begin{figure}[hbt]
\centering{
\includegraphics[width=4.5in]{../newfig/refine.eps}
\caption{Abstraction by reducing latches}
\label{fig:refine}}
\end{figure}
\bi
\item Remove ``irrelevant" latches, reduce state space
\item Provide an over-approximation
\item This algorithm requires \alert{UNSAT core extraction}
\ei
\hyperlink{motiv3}{\beamergotobutton{Recall $k$-BMC}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Refutation within Buchberger's algo.}}
\bi
\item Recall Buchberger's algo....
\pause
\item  What can we learn from the \alert{``1" generated by Buchberger's Algorithm}?
\ei
\begin{Theorem}
\label{thm:weak-ns}
Let $J = \langle f_1,\dots, f_s \rangle $ be
an ideal in the ring $ \mathbb{F}[x_1, \dots, x_d]$ and $V_{\overline{\F}}(J)$ be its
variety over $\overline{\F}$. Then $V_{\overline{\F}}(J) = \emptyset \iff J =
\mathbb{F}[x_1,\dots,x_d] \iff 1 \in J$.  
\end{Theorem}
$V_{\overline{\F}}(J)=\emptyset \iff 1\in J \iff reduced~GB(J) = \{1\}$

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Motivating Example}}
\bi
\item $f_1\sim f_9 \in \mathbb{F}_2[a,b,c,d]$
\item $F = \{f_1,f_2,\dots,f_9\}$
\ei

\begin{minipage}{2in}
\begin{align*}
f_1&: abc + ab + ac + bc \\
   & + a + b + c + 1\\
f_2&: b\\
f_3&: ac\\
f_4&: ac + a
\end{align*}
\end{minipage}
%\hspace{0.1in}
\begin{minipage}{2in}
\begin{align*}
f_5&: bc + c\\
f_6&: abd + ad + bd + d\\
f_7&: cd\\
f_8&: abd + ab + ad + bd + a + b + d + 1\\
f_9&: abd + ab +bd + b
\end{align*}
\end{minipage}
\vspace{1cm}
\bi
\item Find a subset $F_c \subset F$ {\it s.t.} $F_c$ remains UNSAT 
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Motivating Example}}
Following Spoly selection strategy $(f_1,f_2)\to (f_1,f_3) \to (f_2,f_3)\to (f_1,f_4) \to \cdots$,
execute Buchberger's algorithm until adding ``1" to Gr\"obner basis
\bi
\item $Spoly(f_1,f_2)= f_1 - ac\cdot f_2\xrightarrow{a\cdot f_2} ~\xrightarrow{f_3}~
\xrightarrow{c\cdot f_2}  ~\xrightarrow{f_2} f_{10}=a+c+1$
\item $Spoly(f_1,f_3)\xrightarrow{F}_+ 0$
\item \dots
\item $Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$
\item $Spoly(f_2,f_5)\xrightarrow{f_{11}} 1$
\ei
$\{f_1,\dots,f_9,f_{10},f_{11},1\}$ is the Gr\"obner basis generated from Buchberger's algorithm
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Note: this frame contains multiple pages for animation purpose

\begin{frame}{\large{Motivating Example: Refutation Tree}}
\vspace{-0.1in}
\includegraphics[width=3.5in]{./refutation_1.pdf}
\pause
\bi
\item $f_{10}=f_1-acf_2-af_2-f_3-cf_2-f_2 = f_1+acf_2+af_2+f_3+cf_2+f_2$
\item $1 = \F(f_1,f_2,f_3,f_4,f_5)\implies 1 \in \langle f_1,f_2,f_3,f_4,f_5\rangle \implies UNSAT$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Reducing Size: Redundancy in Refutation Tree}}
\bi
\item Expand: \begin{align*}
		1 &= \F(f_1,f_2,f_3,f_4,f_5)\\
		&= cf_2+f_5+\alert{f_{11}}\\
		&=cf_2+f_5+f_3+f_4+\alert{f_{10}}\\
		&=(cf_2 + f_5) + \dots + \mathbf{1\cdot f_3} + \dots + \mathbf{1\cdot
  f_3}+ \dots + (f_1 + acf_2)
		\end{align*}
\item UNSAT core reduced to $\{f_1,f_2,f_4,f_5\}$ which is {\bf minimal}
\item \alert{GB-core algorithm}:
	\bi
	\item Execute Buchberger's algorithm
	\item Recording data including Spoly, polynomials for division and remainder
	\item Terminate Buchberger's algorithm after recording remainder ``1"
	\item Build refutation tree and get UNSAT core
	\item Analyze recorded data, remove redundant polynomials from the core
	\ei
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Reducing Size: Iterative Refinement}}
\bi
\item Spoly pair selection strategy: $(f_1,f_2)\to(f_1,f_3)\to(f_2,f_3)\to(f_1,f_4)\to(f_2,f_4)\to\cdots$
\item High likelihood in minimal core $\to$ Put ahead in Spoly queue $\to$ Faster approaching ``1" in GB-core $\to$ Smaller core
\ei
\vspace{-0.2in}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{refutation_2.pdf}
%\caption{Refutation distance \& frequency}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\bi
\item {\bf Refutation distance}: shortest path to leaf
\item Distance$\downarrow \to$ LT Degree$\downarrow \to$ Likelihood in minimal core $\uparrow$
\vspace{0.2in}
\item {\bf Frequency}: number of times $f_i$ appears in refutation tree
\ei
\end{column}
\end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% use "alignat" to align to multiple columns

\begin{frame}{\large{Iterative Refinement Example}}
\vspace{-0.1in}
\begin{alignat*}{4}
& f_1: x_1x_3+x_3 && ~~~f_2: x_2 + 1 && ~~~f_3: x_2x_3+x_2 \\
& f_4: x_2x_3 && ~~~f_5: x_2x_3 + x_2 + x_3 + 1 && ~~~f_6 : x_1x_2x_3 +x_1x_3
\end{alignat*}
\vspace{-0.5in}
\begin{columns}[onlytextwidth]
\begin{column}{0.6\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.3]{iterative_1.pdf}
\end{figure}
\end{column}
\begin{column}{0.4\textwidth}
\bi
\item $F=\{f_1,f_2,\dots,f_6\}\in\mathbb{F}_2[x_1,x_2,x_3]$
\item UNSAT core: $f_1,f_2,f_3,f_4$
\item \alert{Refutation distance}: \\
		$f_2~~f_1~~f_3~~f_4$\\
		$2~~~3~~~3~~~3$
\item \alert{Frequency}:\\
		$f_1~~f_3~~f_4$\\
		$1~~~1~~~1$
\item Reorder: $\alert{f_2},f_1,f_3,f_4$
\ei
\end{column}
\end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Iterative Refinement Example}}
\vspace{-0.1in}
\begin{alignat*}{4}
& f_1: x_1x_3+x_3 && ~~~f_2: x_2 + 1 && ~~~f_3: x_2x_3+x_2 \\
& f_4: x_2x_3 && ~~~f_5: x_2x_3 + x_2 + x_3 + 1 && ~~~f_6 : x_1x_2x_3 +x_1x_3
\end{alignat*}
\vspace{-0.5in}
\begin{columns}[onlytextwidth]
\begin{column}{0.6\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.3]{iterative_2.pdf}
\end{figure}
\end{column}
\begin{column}{0.4\textwidth}
\bi
\item Initial order: $f_2,f_1,f_3,f_4$
\item Spoly pairs selection:\\
	$(\alert{f_2},f_1)\to(\alert{f_2},f_3)\to(f_1,f_3)\to(\alert{f_2},f_4)\to(f_1,f_4)\to(f_3,f_4)$
\item UNSAT core: $f_2,f_3,f_4$
\item \alert{Fixpoint reached}
\ei
\end{column}
\end{columns}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\large{Reducing Size Further using Syzygy Heuristic}}
\bi
\item Finding interdependencies: $f_i\in\langle F\setminus\{f_i\}\rangle$?
\item Given $F=\{f_1,\dots,f_s\}$, find $f_i = \sum_{j\neq i} h_jf_j$
\item Info lost in Buchberger's algorithm?
\item Inner loop of Buchberger's algorithm: \\
\hspace{0.2in}  $Spoly(f, g) \stackrel{G'}{\textstyle\longrightarrow}_+
r$ \\
\hspace{0.2in}  \alert{IF $r \neq 0$} THEN $G:= G \cup \{r\}$ \\
\hspace{0.2in}  \alert{IF $r = 0$} THEN Discard

\item We collect division info when $r = 0$:
	\bi
	\item Record data for Spoly and polynomial division as in GB-core algorithm
	\ei
\item $Spoly(f_i,f_j)\xrightarrow{F}_+ 0\implies \alert{c_1f_1+c_2f_2+\cdots+c_sf_s = 0}$
\pause
\item $(c_1,c_2,\dots,c_s)$ is a {\bf syzygy} on $(f_1,f_2,\dots,f_s)$
\ei
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Reducing Size Further using Syzygy Heuristic}}
\bi
\item In a single syzygy, $c_i=1 \implies f_i = \sum_{j\neq i} h_jf_j$
\item In general cases, need to analyze all syzygies recorded
\item Collect $m$ syzygies as a system of polynomial equations, or a {\bf Syzygy Matrix}
\ei
\hspace{-0.1in}
\begin{minipage}{2.5in}
\begin{equation*} \label{eqn:syz}
 \begin{cases}
 c_1^1f_1+c_2^1f_2+\cdots+c_s^1f_s = 0\\
 c_1^2f_1+c_2^2f_2+\cdots+c_s^2f_s  = 0\\
 \ \ \ \ \ \  \vdots \\
 c_1^mf_1+c_2^mf_2+\cdots+c_s^mf_s = 0  
 \end{cases}
\end{equation*}
\end{minipage}
\hspace{-0.3in}
\begin{minipage}{2in}
%which can be represented in matrix form as:
\begin{center}
\begin{align*}
\label{mat:syzygy}
   \begin{bmatrix}
           c_1^1 & c_2^1 & \cdots & c_s^1 \\
           c_1^2 & c_2^2 & \cdots & c_s^2 \\
           \vdots & \vdots & \ddots & \vdots \\
           c_1^m & c_2^m & \cdots & c_s^m
         \end{bmatrix}
    \begin{bmatrix}
           f_{1} \\
           f_{2} \\
           \vdots \\
           f_{s}
         \end{bmatrix}
         &= 0
  \end{align*}

\end{center}
\end{minipage}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Reducing Size: Revisiting Motivating Example with Syzygy}}
\vspace{-0.2in}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{refutation_tree.pdf}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\bi
\item Recorded in Buchberger's Algo:\\
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$\\
$Spoly(f_3,f_4)\xrightarrow{F}_+ f_{11}$\\
$Spoly(f_2,f_5)\xrightarrow{F}_+ 1$
\item Discarded in Buchberger's Algo:\\
$Spoly(f_1,f_3)\xrightarrow{F}_+ 0$\\
$Spoly(f_2,f_3)\xrightarrow{F}_+ 0$\\
$\dots$\\
$Spoly(f_1,f_5)\xrightarrow{F}_+ 0$
\ei
\end{column}
\end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Syzygy Example}}
\bi
\item Syzygy matrix:
\ei
\includegraphics[width=4.5in]{./syzygy_1.pdf}
\pause
\llap{\includegraphics[width=4.5in]{./syzygy_2.pdf}}
\bi
\item $J = \langle f_1,f_2,\dots,f_9\rangle$
\item $f_{10}\in J$
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Syzygy Example}}
\bi
\item Considering $f_{10}=f_1-acf_2-af_2-f_3-cf_2-f_2 = f_1+acf_2+af_2+f_3+cf_2+f_2$
\ei
\includegraphics[width=4.5in]{./syzygy_3.pdf}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Syzygy Example}}
\includegraphics[width=4.5in]{./syzygy_4.pdf}
\bi
\item Column $f_3$ contains ``1", means:\\
$f_3 = ac f_2 + a f_5$
\item Generalization of this strategy refer to the paper
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Overall Approach}}
\begin{algorithm}[H] % note [hbt] will fail in beamer
\SetAlgoNoLine
 \KwIn{A set of UNSAT polynomials $F$} 
 \KwOut{A subset $F'\subset F$ remains UNSAT}
 $G \gets F$\;
  \Repeat{$G=F'$}
  {
	$F' = G$\;
  	$G \gets $GB-core($F'$ with order $>$)\;
	Update order $>$\;
  }
  $F'\gets$syzygy\_heuristic($G$)\;
\Return{$F'$}
\caption {UNSAT core extraction based on Gr\"obner basis algorithm}
\end{algorithm}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Experiment results: UNSAT core extraction}}
\centering{Table of selected benchmarks}\\
\centering{I:single GB-core; II: Iterative GB-core; III: syzygy}
\includegraphics[width=4.5in]{./table_cp2016.pdf}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Application to abstraction refinement}}
\vspace{-0.1in}
\begin{columns}[onlytextwidth]
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{../newfig/s27_stg.pdf}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
\centering
\includegraphics[scale=0.25]{../newfig/s27_stg_refined.pdf}
\end{figure}
\end{column}
\end{columns}
\bi
\item By removing ``irrelevant" latches, state-space reduced
\ei
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Conclusion}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Future work}}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}{\large{Publications \& tools}}
\end{frame}

\end{document}